### 2013-06-26 LNCS_latex_template.tex
alles eingechecked

%%%%%%%%%%%%%%%%%%%%%%% file typeinst.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is the LaTeX source for the instructions to authors using
% the LaTeX document class 'llncs.cls' for contributions to
% the Lecture Notes in Computer Sciences series.
% http://www.springer.com/lncs       Springer Heidelberg 2006/05/04
%
% It may be used as a template for your own input - copy it
% to a new file with a new name and use it as the basis
% for your article.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[runningheads]{llncs}

\usepackage{amssymb}
\setcounter{tocdepth}{3}
\usepackage{graphicx}
\usepackage{tabularx}

\usepackage{url}
\urldef{\mailsa}\path |c.slawtschew@tu-berlin.de|
\urldef{\mailsb}\path |c.slawtschew@tu-berlin.de|
\newcommand{\keywords}[1]{\par\addvspace\baselineskip
\noindent\keywordname\enspace\ignorespaces#1}

\begin{document}

\mainmatter  % start of an individual contribution

% first the title is needed
\title{The Title ...}

% a short form should be given in case it is too long for the running head
\titlerunning{Lecture Notes in Computer Science: Authors' Instructions}

% the name(s) of the author(s) follow(s) next
%
% NB: Chinese authors should write their first names(s) in front of
% their surnames. This ensures that the names appear correctly in
% the running heads and the author index.
%
\author{Author}
%
\authorrunning{Author}   % abbreviated author list (for running head)
%
%%%% modified list of authors for the TOC (add the affiliations)
\tocauthor{Author (Berlin Institute of Technology)}
%
\institute{Berlin Institute of Technology\\Chair of Systems Analysis and Computing\\Franklinstr. 28/29, 10587 Berlin, Germany\\
\email{emailadress}\\ WWW home page:
\texttt{http://www.sysedv.tu-berlin.de}}
%
% NB: a more complex sample for affiliations and the mapping to the
% corresponding authors can be found in the file "llncs.dem"
% (search for the string "\mainmatter" where a contribution starts).
% "llncs.dem" accompanies the document class "llncs.cls".
%

\toctitle{Lecture Notes in Computer Science}
\tocauthor{Authors' Instructions}
\maketitle

\begin{abstract} abstract abstract bla bla ...

 \keywords{We would like to encourage you to list your keywords within
the abstract section}
\end{abstract}

\section{Introduction}

...

\section{Related Work}
There have been active studies related to the problem in both
acdemia and industry…
\\
There are some studies, mostly from academia, done for …
\\
Some previous work envisioned the task of
\\
We believe that this is an interesting approach, in ist own right,
but we do not discuss this approach further in this paper.

\section{Section 3}

\section{Section 4}

\section{Concluding Remarks}
...
\\

Please use the commands \verb+\label+ and \verb+\ref+ for
cross-references and the commands \verb+\bibitem+ and \verb+\cite+ for
references to the bibliography, to enable us to create hyperlinks at
these places.

For preparing your figures electronically and integrating them into
your source file we recommend using the standard \LaTeX{} \verb+graphics+ or
\verb+graphicx+ package. These provide the \verb+\includegraphics+ command.
In general, please refrain from using the \verb+\special+ command.

\subsubsection{Lemmas, Propositions, and Theorems.}

The numbers accorded to lemmas, propositions, and theorems, etc. should
appear in consecutive order, starting with Lemma 1, and not, for
example, with Lemma 11.

\subsection{Figures}

For \LaTeX\ users, we recommend using the \emph{graphics} or \emph{graphicx}
package and the \verb+\includegraphics+ command.

Please check that the lines in line drawings are not
interrupted and are of a constant width. Grids and details within the
figures must be clearly legible and may not be written one on top of
the other. Line drawings should have a resolution of at least 800 dpi
(preferably 1200 dpi). The lettering in figures should have a height of
2~mm (10-point type). Figures should be numbered and should have a
caption which should always be positioned \emph{under} the figures, in
contrast to the caption belonging to a table, which should always appear
\emph{above} the table; this is simply achieved as matter of sequence in
your source.

Please center the figures or your tabular material by using the \verb+\centering+
declaration. Short captions are centered by default between the margins
and typeset in 9-point type (Fig.~\ref{fig:example} shows an example).
The distance between text and figure is preset to be about 8~mm, the
distance between figure and caption about 6~mm.

To ensure that the reproduction of your illustrations is of a reasonable
quality, we advise against the use of shading. The contrast should be as
pronounced as possible.

If screenshots are necessary, please make sure that you are happy with
the print quality before you send the files.
\begin{figure}
\centering
%\includegraphics[height=6.2cm]{eijkel2}
\caption{One kernel at $x_s$ (\emph{dotted kernel}) or two kernels at
$x_i$ and $x_j$ (\textit{left and right}) lead to the same summed estimate
at $x_s$. This shows a figure consisting of different types of
lines. Elements of the figure described in the caption should be set in
italics, in parentheses, as shown in this sample caption}
\label{fig:example}
\end{figure}

Please define figures (and tables) as floating objects. Please avoid
using optional location parameters like ``\verb+[h]+" for ``here".

\paragraph{Remark 1.}

In the printed volumes, illustrations are generally black and white
(halftones), and only in exceptional cases, and if the author is
prepared to cover the extra cost for color reproduction, are colored
pictures accepted. Colored pictures are welcome in the electronic
version free of charge. If you send colored figures that are to be
printed in black and white, please make sure that they really are
legible in black and white. Some colors as well as the contrast of
converted colors show up very poorly when printed in black and white.

\subsection{Formulas}

Displayed equations or formulas are centered and set on a separate
line (with an extra line or halfline space above and below). Displayed
expressions should be numbered for reference. The numbers should be
consecutive within each section or within the contribution,
with numbers enclosed in parentheses and set on the right margin --
which is the default if you use the \emph{equation} environment, e.g.
\begin{equation}
  \psi (u) = \int_{o}^{T} \left[\frac{1}{2}
  \left(\Lambda_{o}^{-1} u,u\right) + N^{\ast} (-u)\right] dt \;  .
\end{equation}

Please punctuate a displayed equation in the same way as ordinary
text but with a small space before the end punctuation.

\subsection{Footnotes}

The superscript numeral used to refer to a footnote appears in the text
either directly after the word to be discussed or -- in relation to a
phrase or a sentence -- following the punctuation sign (comma,
semicolon, or period). Footnotes should appear at the bottom of
the normal text area, with a line of about 2~cm set
immediately above them.\footnote{The footnote numeral is set flush left
and the text follows with the usual word spacing.}

\subsection{Program Code}

Program listings or program commands in the text are normally set in
typewriter font, e.g., CMTT10 or Courier.

\medskip

\noindent
{\it Example of a Computer Program}
\begin{verbatim}
program Inflation (Output)
  {Assuming annual inflation rates of 7%, 8%, and 10%,...
   years};
   const
     MaxYears = 10;
   var
     Year: 0..MaxYears;
     Factor1, Factor2, Factor3: Real;
   begin
     Year := 0;
     Factor1 := 1.0; Factor2 := 1.0; Factor3 := 1.0;
     WriteLn('Year  7% 8% 10%'); WriteLn;
     repeat
       Year := Year + 1;
       Factor1 := Factor1 * 1.07;
       Factor2 := Factor2 * 1.08;
       Factor3 := Factor3 * 1.10;
       WriteLn(Year:5,Factor1:7:3,Factor2:7:3,Factor3:7:3)
     until Year = MaxYears
end.
\end{verbatim}
%
\noindent
{\small (Example from Jensen K., Wirth N. (1991) Pascal user manual and
report. Springer, New York)}

\subsection{Citations}

The list of references is headed ``References" and is not assigned a
number. The list should be set in small print
and placed at the end of your contribution, in front of the appendix,
if one exists.
Please do not insert a pagebreak before the list of references if the
page is not completely filled.
An example is given at the end of this information sheet. For citations in the text you can set square brackets and consecutive numbers: \cite{leeuw},
\cite{bru:car:pier}, \cite{mich} \newline
But you could also choose other bibliography-styles.
If you are using bibtex for your bibliography, you can specify the style easily by using the command \verb+\bibliographystyle{}+ after your bibliography command at the end of the document. Between the brackets you specify the kind of style you prefer. In LaTex there are several styles possible ,for example \verb+apalike+, \verb+\plaindin+ or \verb+\dinat+.\newline
For more details for right citation and citation-styles please see also the chair guidelines for scientific papers, available on the website: \newline \verb+http://www.sysedv.tu-berlin.de+\dots

\subsection{Page Numbering and Running Heads}

Please set page numbers by citing within the text. For example, after a written section\cite[p. 5.]{leeuw}

\section{LNCS Online}

The online version of the volume will be available in LNCS Online.
Members of institutes subscribing to the Lecture Notes in Computer
Science series have access to all the pdfs of all the online
publications. Non-subscribers can only read as far as the abstracts. If
they try to go beyond this point, they are automatically asked, whether
they would like to order the pdf, and are given instructions as to how
to do so.

\section{BibTeX Entries}

The correct BibTeX entries for the Lecture Notes in Computer Science
volumes can be found at the following Website shortly after the
publication of the book:
\url{http://www.informatik.uni-trier.de/~ley/db/journals/lncs.html}

\section{Lecture Notes in Computer Science in the ISI SCI Expanded}

The Lecture Notes in Computer Science volumes are sent to ISI for
inclusion in their Science Citation Index Expanded.

\subsubsection*{Acknowledgments.} The heading should be treated as a
subsubsection heading and should not be assigned a number.

\begin{thebibliography}{4}
%

%
\bibitem{leeuw}
van Leeuwen, J. (ed.):
Computer Science Today. Recent Trends and Developments.
Lecture Notes in Computer Science, Vol.~1000.
Springer-Verlag, Berlin Heidelberg New York (1995)
%
\bibitem{bru:car:pier}
Bruce, K.B., Cardelli, L., Pierce, B.C.:
Comparing Object Encodings.
In: Abadi,~M., Ito,~T. (eds.):
Theoretical Aspects of Computer Software.
Lecture Notes in Computer Science, Vol.~1281.
Springer-Verlag, Berlin Heidelberg New York (1997) 415--438
%
\bibitem{mich}
Michalewicz, Z.:
Genetic Algorithms + Data Structures = Evolution Programs.
3rd edn. Springer-Verlag, Berlin Heidelberg New York (1996)
%
\bibitem{bal:cha:gra:pae}
Baldonado, M., Chang, C.-C.K., Gravano, L., Paepcke, A.:
The Stanford Digital Library Metadata Architecture.
Int. J. Digit. Libr. \textbf{1} (1997) 108--121
%
\end{thebibliography}

\section*{Appendix: Springer-Author Discount}

LNCS authors are entitled to a 33.3\% discount off all Springer
publications. Before placing an order, the author should send an email
to \texttt{SDC.bookorder@springer.com}, giving full
details of his or her
Springer publication, to obtain a so-called token. This token is a
number, which must be entered when placing an order via the Internet, in
order to obtain the discount.

\section{Checklist of Items to be Sent to Volume Editors}
Here is a checklist of everything the volume editor requires from you:

\begin{itemize}
\settowidth{\leftmargin}{{\Large$\square$}}\advance\leftmargin\labelsep
\itemsep8pt\relax
\renewcommand\labelitemi{{\lower1.5pt\hbox{\Large$\square$}}}

\item The final \LaTeX{} source files
\item A final PDF file
\item A copyright form, signed by one author on behalf of all the
authors of the paper.
\item A readme giving the first name(s) and the surname(s) of all the
authors of the paper, as well as the name and address of the
corresponding author.
\end{itemize}
\end{document}

### 2013-06-26 paper.tex
alles eingechecked

\documentclass[runningheads]{llncs}

\usepackage[utf8]{inputenc}
\usepackage[numbers, comma, sort]{natbib}
\bibliographystyle{apalike}

\usepackage{amssymb}
\setcounter{tocdepth}{3}
\usepackage{graphicx}
\usepackage{float}

\usepackage{pgfplots}
\usepackage{tikz}
\usetikzlibrary{arrows,backgrounds,positioning}

\usepackage{url}
\urldef{\mailsa}\path|rob@clabs.cc, schubotz@tu-berlin.de|
\newcommand{\todaydate}{\leadingzero{\day}.\leadingzero{\month}.\the\year} 
\newcommand{\keywords}[1]{\par\addvspace\baselineskip
\noindent\keywordname\enspace\ignorespaces#1}

\begin{document}

\mainmatter

\title{Mathematical Language Processing \\ Project}
\titlerunning{MLP}

\author{Robert Pagel \and Moritz Schubotz}
\authorrunning{Pagel, Schubotz}

\institute{Berlin Institute of Technology, Database Systems and Information Management Group,\\
Einsteinufer 17, 10587 Berlin, Germany\\
\mailsa\\
\url{http://www.dima.tu-berlin.de/}}

\maketitle

\begin{abstract}
The \emph{Mathematical Language Processing project} addresses the problem of discovering definition relations between variable symbols and the surrounding text. For machine learning approaches like support vector machines, a labelled text corpus is a prerequisite. Unlike traditional approaches, we propose a strategy for automated data labelling with very limited supervision. We analyse the surface of the surrounding text and track co-occurrences of artifacts within the sentences to extract relations. 
%\keywords{machine learning, text mining, parallel computing}
\end{abstract}

\section{Introduction}

\subsection{Motivation}
Mathematical formulas are a viable source of information for a wide range of scientists. Often those formulas contain variables that are unknown or at least ambiguous to the reader. Therefore, one needs to study the surrounding text to find the relevant definition.

An automatic information retrieval system can be used to reduce the readers effort to understand a formula. Especially students and scientists of other disciplines would profit of such a system.

Unfortunately, we were not able to find a labelled text corpus that annotates variables and their definition. Instead of labelling large amounts of text by hand, we investigate a potential way of automatically labelling large corpora with a set of statistical assumptions. One can derive distributional statistics of the found patterns on a large labelled corpus. This statistics may indicate \emph{overfitting features} and, therefore, help to mitigate dimensionality issues with a classifier \cite{Oommen2008}.

We chose the Wikipedia text corpus as a target because of two facts. First, the mark-up for editing formulas is a subset of \TeX. Therefore, it enables us to collect all variables within the set of formulas in a document. Second, we can leverage the semantic information between mark-up and word tokens, e.g., links to other articles.

The English Wikipedia contains roughly four million articles. Even if we only pick articles containing \texttt{<math/>} mark-up, our processor still needs to compute tens of thousands of articles. Especially when using a maximum entropy \emph{POS} tagger \cite{Rathna96}, like the one in Stanford's NLP framework, one can make use of a parallel processing system to speed up computation.

We will show how the proposed strategy can be implemented in the PACT programming model \cite{Alexandrov2010}.

\subsection{Related Work}
\citeauthor{Quoc2010} (2010) proposed in \cite{Quoc2010} an approach for relating whole formulas to sentences and their describing paragraphs. \citeauthor{Yokoi} (2010) trained a \emph{support vector machine} to extract natural language descriptions for mathematical expressions \cite{Yokoi}.

\section{Relation Discovery}
The basic assumption of our approach is that if two entities take part in a relation, those two co-occurre in the same sentence. This assumption is also used in the \emph{Distant Supervision} \cite{Mintz2008} approach by \citeauthor{Mintz2008} Furthermore, we assume in a scientific document, which contains a specific variable name, there is at least one sentence with such a relation.

The two entities have some characteristics that can be used for filtering. First of all, the relevant variables can be retrieved accurately from \TeX\ formatted formulas (see Section \ref{vr}). The fact that a definition relation consists of two entities and we know that one must be the variable, leads to an obvious conclusion: Sentences without a variable cannot contain a definition relation and can be discarded.
Second, one can assume that the other entity must be a noun or a compound noun. Thus, all words that are not nouns or adjectives describing these nouns, can also be discarded.

\subsection{Numerical Statistics}
As a prerequisite for detecting potential relations, one needs to identify all occurrences of a variable within a document. Our intuition tells us, that the corresponding definition relation entities can be found near the variables. In other words, the distance between a variable and an entity correlates to the probability of being a definition relation.

We translate this into a rating function based on the gaussian curve that takes the entity word positions $p_{1}$ and $p_{2}$ and relates the resulting distance to a value in $[0;1]$.

\begin{equation}
\label{eq:dist}
R_{d}(p_{1}-p_{2}) = R_{d}(d) = \exp\left[-\frac{d^{2}}{2c_{d}^{2}}\right]
\end{equation}

The parameter $c_{d}$ relates to the full width at half maximum. We assume that most of the definition relations span across five words. One might choose $c_{s}$, such that $R_{d}(p_{1},p_{2}) = \frac{1}{2}$ if $\left|p_{2}-p_{1}\right| = 5$. Therefore, $c_{d} = 5 / \sqrt{2\log{2}}$. Please note that this assumption is heavily influenced by the language of the text corpus.

Not only the distances within one sentence are of interest. Also the sequential order of sentences, containing a variable, carry some viable sources for statistics. Usually, we experiance that an author first introduces a term and uses it in subsequent sentences without necessarily repeating its definition. We assume that this kind of behaviour also applies to the use of variables. Thus, we extend the idea of distance to the sequential order of sentences. Here, the distance is the count $n$ of candidate sentences, beginning at the top of the document.

\begin{equation}
\label{eq:order}
R_{s}(n) = \exp\left[-\frac{n^{2}}{2c_{s}^{2}}\right]
\end{equation}

We define the parameter $c_{s}$ analogous to $c_{d}$ used in Equation \ref{eq:dist}. In our experiments, we expect that most of the definition relations are found within the first three sentences. Thus, $c_{s} = 3 / \sqrt{2\log{2}}$.

\begin{figure}[H]
\centering
\begin{tikzpicture}
	\begin{axis}[
			axis x line = bottom,
			axis y line = left,
			height=3.5cm, width=12cm,
			xmin=-11, xmax=11,
			ymin=0, ymax=1,
			xtick={-5,0,5},
			ytick={0,0.5,1},
			xlabel=$Distance$,
			ylabel=$Ranking$]
		\addplot [mark=,domain=-20:20,smooth]{e^(-( x^2 / 36.06737602222409 ))};
		\addplot [mark=,domain=-5:5,smooth, gray,fill=gray!90!black,opacity=0.5]{e^(-( x^2 / 36.06737602222409 ))} \closedcycle;
		\addplot [black,dotted] coordinates {(0,0) (0,1)};
		\addplot [black,dashed] coordinates {(5,0) (5,1)};
		\addplot [black,dashed] coordinates {(-5,0) (-5,1)};
	\end{axis}
\end{tikzpicture}
\caption{Curve of the rating function $R_{d}$ with FWHM}
\end{figure}

Furthermore, we take distributional properties into account. The classic tf-idf \cite{Salton86} statistic reflects the importance of a term to a document. Unfortunately, the \emph{inverse document frequency} is not applicable to definition relations. In our approach, the frequency distribution of all terms, that participate in various definition relations is unknown beforehand. Thus, using tf-idf could affect the sensitivity of our rating statistic significantly. E.g., `Hamiltonian' as well as `length' are both valid entities. Obviously, the number of documents (respectively sentences) containing one of these words, within a corpus like Wikipedia, differs a lot.

Therefore, we use the normalized term frequency. $t$ and $w$ are terms. $f$ denotes the raw frequency of a term within a document $d$.

\begin{equation}
	\label{eq:tf}
	\mathrm{tf}(t,d) = \frac{\mathrm{f}(t,d)}{\max\{\mathrm{f}(w,d):w \in d\}}
\end{equation}

We define our ranking statistic as the weighted mean of the rankings defined in Equations \ref{eq:dist}, \ref{eq:order}, and \ref{eq:tf}:

\begin{equation}
	\label{eq:rating}
	\frac{\alpha\cdot\mathrm{R}_{d}+\beta\cdot\mathrm{R}_{s}+\gamma\cdot\mathrm{tf}}{\alpha+\beta+\gamma} \mapsto [0;1]
\end{equation}

\section{Implementation}

\subsection{Overview of the processor}
The MLP processing system \cite{github} is a graph of PACTs, where data flows through. In contrast to other programming models like Map/Reduce, PACTs can operate on multiple inputs. At some point of time, we need an ordered list of tokenized sentences and a list of identified variables, to compute the ranking score. A \emph{CoGroup Contract} takes both lists and computes the subsets, build upon distinct document keys. Finally, a \emph{Reduce Contract} filters out results below a given threshold $\theta$.

\tikzset{actor/.style={
        rectangle,
        minimum size=6mm,
        very thick,
        draw=gray!50!black!50,
        top color=white,
        bottom color=gray!50!black!20
    },
    arrow/.style={
        -latex, thick, shorten <=2pt,shorten >=2pt
    }
}
\begin{figure}[H]
	\centering
	\begin{tikzpicture}[node distance=5mm and 8mm]
		\node (Input) [align=center]{Wiki Dumps};
		\node (DocumentParser) [actor, right=of Input, align=center] {\emph{Map}\\\textbf{Parser}};
		\node (Candidates) [actor, right=of DocumentParser, align=center] {\emph{CoGroup}\\\textbf{Kernel}};
		\node (Sentence) [actor, above=of DocumentParser, align=center] {\emph{Map}\\\textbf{Tagger}};
		\node (Filter) [actor, right=of Candidates, align=center] {\emph{Reduce}\\\textbf{Filter}};
		\node (Output) [right=of Filter,align=center]{Relation\ Candidates};
		\draw[arrow] (Input)--(DocumentParser);
		\draw[arrow] (DocumentParser)--(Sentence);
		\draw[arrow] (DocumentParser)--(Candidates);
		\draw[arrow] (Sentence.east)--(Candidates.north);
		\draw[arrow] (Candidates)--(Filter);
		\draw[arrow] (Filter)--(Output);
	\end{tikzpicture}
\caption{Data flow of the PACT program}
\end{figure}
There are three computation heavy parts in the processor. First, the raw Wikipedia document dumps need to be transformed into plain text. At the same step, the formulas need to be evaluated and replaced by place-holder words. We used the \emph{WikiText} framework of the Mylyn project with a custom builder to transform the document into plain text. For analysing the \TeX \ mark-up in the \texttt{<math/>} tags, \emph{SnuggleTeX} transforms the formulas into their MathML representations, where variable names can be extracted easily. The second computation heavy part is tokenization and tagging of the plain text documents. We used the Stanford maximum entropy tagger with the \emph{WSJ left3words} language model. Finally, the rating for all nouns, contained in all candidate sentences, needs to be calculated.

\subsection{Experiments}

\subsubsection{Variable Retrieval}
\label{vr}
Throughout our experiments we made some observations that had an impact on the accuracy of retrieving the correct set of variables. First of all, people tend to incorrectly use \TeX\ trying to create formulas. I.e., \texttt{\textbackslash text\{log\}} is more often used than the correct operator \texttt{\textbackslash log}. Another problem is that sometimes people use indices as a form of annotation, like $T_{before}$, $R_{A}$, and $R_{B}$. These ambiguities lead to a increased number of false positives.

We took a very conservative approach and preprocessed all formulas. \texttt{\textbackslash text\{\}} blocks along with plain subscriptions will be removed before analysis. Moreover, we created a comprehensive blacklist to improve the results further. Short words like `a' and `i', which are very common, could be easily matched by our processor if a formula contains one of those symbols. Additionally, we blacklist common mathematical operators, constants, and functions.

Because of the fact, that we do not have any annotated test corpora, evaluation has to be performed by hand. Therefore, we took a sample of 30 random documents and counted all matches. The resulting estimates for \emph{recall} and \emph{precision} are 0.99 and 0.86 respectively.

\subsubsection{Tuning}
The weighting parameters $\alpha$, $\beta$, and $\gamma$ in Equation \ref{eq:rating} are used to balance the contribution of all embedded statistics to the overall ranking. But these parameters need to be tuned carefully. Otherwise the results tend to be biased towards one parameter. In other words, if the distance measure $R_{D}$ dominates the other measures, more complex\footnote{in terms of words between both entities, participating in a relation} relations are less likely to be discovered.

The \emph{term frequency} turns out to be the least robust statistic. Usually, the frequency of a specific term co-occurring with a variable name within one document is rather low. Thus, the derived ranking is relatively volatile. In our experiments, we chose $\gamma$ to be 0.75 ($\alpha = \beta = 1$) for a more robust ranking.

\section{Conclusion}
Our experiments showed that combining a POS tagger with numerical statistics about the text surface, can lead to quality results. However, this approach is only applicable under certain conditions. In situations where both parts of a relation are unknown, other methods, especially supervised ones, are better used to bootstrap a language model. Unfortunately, we still need an other labelled test corpus to measure the real performance of a classifier, trained with the derived features.

Apart from the mentioned weighting parameters, the ranking's quality is influenced by two further parameters. These are the assumptions about the parameters $c_{d}$ and $c_{s}$. Balancing all these parameters requires extensive knowledge about the underlying language. These parameters might also be specific for every type of relation. Therefore, this strategy depends heavily on human input and many iterations of tuning.

To further improve the robustness of the \emph{term frequency} measure, one might cluster sets of scientific documents, based on their specific field of research.

\bibliography{mlp-papers}
\end{document}

### 2013-06-26 paper.tex
Little changes to abstract and some corrections to calculations.

@@
 \begin{abstract}
-The \emph{Mathematical Language Processing project} addresses the problem of discovering definition relations between variable symbols and the surrounding text.
-For machine learning approaches like support vector machines, a labelled text corpus is a prerequisite.
-Unlike traditional approaches, we propose a strategy for automated data labelling with very limited supervision.
-We analyse the surface of the surrounding text and track co-occurrences of artifacts within the sentences to extract relations.
-%\keywords{machine learning, text mining, parallel computing}
+In this paper, we present a solution for automated definition discovery for the variables occurring in mathematical formulae.
+Ordinary natural language processing approaches that are trained based on large manually annotated news corpora fall short in reasonable quality of definition discovery for scientific texts.
+In the \emph{Mathematical Language Processing} project we discover definition relations between variables and the surrounding text with a machine learning approach for classification.
+We extract the identifiers from the formulae and track co-occurrences of those in the surrounding text.
+The evaluation of our prototypical system, applied on the Wikipedia text corpus shows, that our approach augments the user experience for articles with mathematical formula substantially.
+While hovering the identifiers occurring in the equation a pop-up with the most probable explanations for the identifiers occurs.
+The displayed definition information provides a good match to the actual meaning of the identifiers.
+\keywords{machine learning, text mining, parallel computing}
 \end{abstract}
@@
 \subsection{Motivation}
-Mathematical formulas are a viable source of information for a wide range of scientists.
+Mathematical formulae are a viable source of information for a wide range of scientists.
 Often those formulas contain variables that are unknown or at least ambiguous to the reader.
@@
 
-An automatic information retrieval system can be used to reduce the readers effort to understand a formula. Especially students and scientists of other disciplines would profit of such a system.
+An automatic information retrieval system can be used to reduce the readers effort to understand a formula. Especially students and scientists of other disciplines would profit from such a system.
 
-Unfortunately, we were not able to find a labelled text corpus that annotates variables and their definition.
+Unfortunately, no one has built a labelled text corpus that annotates variables and their definition.
 Instead of labelling large amounts of text by hand, we investigate a potential way of automatically labelling large corpora with a set of statistical assumptions.
-One can derive distributional statistics of the found patterns on a large labelled corpus.
-This statistics may indicate \emph{overfitting features} and, therefore, help to mitigate dimensionality issues with a classifier \cite{Oommen2008}.
 
@@
 Therefore, it enables us to collect all variables within the set of formulas in a document.
-Second, we can leverage the semantic information between mark-up and word tokens, e.g., links to other articles.
+Second, we can leverage the semantic information between mark-up and word tokens, e.g., hyperlinks.
 
@@
 
-\section{Relation Discovery}
-The basic assumption of our approach is that if two entities take part in a relation, those two co-occurre in the same sentence.
-This assumption is also used in the \emph{Distant Supervision} \cite{Mintz2008} approach by \citeauthor{Mintz2008} Furthermore, we assume in a scientific document, which contains a specific variable name, there is at least one sentence with such a relation.
+\section{Relation Detection}
+The basic assumption of our approach is that if two entities take part in a relation, those two co-occur in the same sentence.
+This assumption is also used in the \emph{Distant Supervision} \cite{Mintz2008} approach by \citeauthor{Mintz2008}.
+Furthermore, we assume in a scientific document, which contains a specific variable name, there is at least one sentence with such a relation.
 
@@
 First of all, the relevant variables can be retrieved accurately from \TeX\ formatted formulas (see Section \ref{vr}).
-The fact that a definition relation consists of two entities and we know that one must be the variable, leads to an obvious conclusion: Sentences without a variable cannot contain a definition relation and can be discarded.
+The fact that a relation consists of two entities and we know that one must be the variable, leads to an obvious conclusion: Sentences without a variable cannot contain a definition relation and can be discarded.
 Second, one can assume that the other entity must be a noun or a compound noun. Thus, all words that are not nouns or adjectives describing these nouns, can also be discarded.
@@
 As a prerequisite for detecting potential relations, one needs to identify all occurrences of a variable within a document.
-Our intuition tells us, that the corresponding definition relation entities can be found near the variables.
-In other words, the distance between a variable and an entity correlates to the probability of being a definition relation.
+We assume that the distance between the identifier and the corresponding definition is normal distributed.
 
-We translate this into a rating function based on the gaussian curve that takes the entity word positions $p_{1}$ and $p_{2}$ and relates the resulting distance to a value in $[0;1]$.
+Formally, we calculate the distance in words $\Delta=p_\mathrm{def}-p_\mathrm{id}$  between $p_\mathrm{def}$ the position of the definition and $p_\mathrm{id}$ the position of the identifier.
 
+Thus, the assumed normalized relation probability density function reads
 \begin{equation}
 \label{eq:dist}
-R_{d}(p_{1}-p_{2}) = R_{d}(d) = \exp\left[-\frac{d^{2}}{2c_{d}^{2}}\right]
+R_\sigma(\Delta)= {\sigma \sqrt{2\pi}}{e^{{\left(2\sigma^2\right)^{-1}}}} \mathcal N(0,\sigma^2)(\Delta)\equiv \exp\left[-\frac{1}{2}\frac{\Delta^2-1}{\sigma^2}\right].
 \end{equation}
+\newcommand{\normalPlot}[2]{(e^((-1/2)*(((#1)^2-1)/(#2)^2)))} %1/((#2)*sqrt(2*pi)))*
+\newcommand{\normalPlotFWHM}[2]{\normalPlot{#1}{ sqrt(((#2)^2-1)/(2*ln(2)))}}
 
-The parameter $c_{d}$ relates to the full width at half maximum.
-We assume that most of the definition relations span across five words.
-One might choose $c_{s}$, such that $R_{d}(p_{1},p_{2}) = \frac{1}{2}$ if $\left|p_{2}-p_{1}\right| = 5$.
-Therefore, $c_{d} = 5 / \sqrt{2\log{2}}$.
+Thereby we performed the normalisation, with regard to the model assumption that the probability to find a relation at $\Delta=1$ is maximal, i.e. $R(1)=1$. 
+Manual evaluation of some articles indicated that $R(1)\approx\frac{1}{2}R(5)$.
+As, the full width half maximum is at $\Delta=\sqrt{2 \ln(2)  \sigma^2+1}$, and the standard deviation evaluates to $\sigma(\Delta)=\sqrt\frac{\Delta^2-1}{2\ln 2}$.
+As a consequence we choose $\sigma_d=\sqrt\frac{5^2-1}{2\ln 2}$ and write $R_\mathrm{d}(\Delta) \equiv R_{\sigma_\mathrm{d}}(\Delta) $.
 Please note that this assumption is heavily influenced by the language of the text corpus.
@@
 Also the sequential order of sentences, containing a variable, carry some viable sources for statistics.
-Usually, we experiance that an author first introduces a term and uses it in subsequent sentences without necessarily repeating its definition.
+Typically an author first introduces a term and uses it in subsequent sentences without necessarily repeating its definition.
 We assume that this kind of behaviour also applies to the use of variables.
@@
 
+In our experiments, we expect that half of the definition relations are found in the sentence with 3 sentences difference. Thus we choose 
 \begin{equation}
 \label{eq:order}
-R_{s}(n) = \exp\left[-\frac{n^{2}}{2c_{s}^{2}}\right]
+R_{s}(n)={\sigma_\mathrm{s} \sqrt{2\pi}} \mathcal N(0,\sigma_\mathrm{s}^2)(n)= \exp\left[-\frac{1}{2}\frac{n^2}{\sigma_\mathrm{s}^2}\right]
 \end{equation}
-
-We define the parameter $c_{s}$ analogous to $c_{d}$ used in Equation \ref{eq:dist}.
-In our experiments, we expect that most of the definition relations are found within the first three sentences.
-Thus, $c_{s} = 3 / \sqrt{2\log{2}}$.
+ with $\sigma_\mathrm{s}=\frac{3}{\sqrt{2\ln 2}}$. 
+\newcommand{\normalPlotFWHMII}[2]{(e^((-1/2)*(((#1)^2*2*ln(2))/(#2)^2)))}
 
 \begin{figure}[H]
-\centering
 \begin{tikzpicture}
@@
 			height=3.5cm, width=12cm,
-			xmin=-11, xmax=11,
+			xmin=-10, xmax=10,
 			ymin=0, ymax=1,
-			xtick={-5,0,5},
+			xtick={-5,-3,0,3,5},
 			ytick={0,0.5,1},
 			xlabel=$Distance$,
-			ylabel=$Ranking$]
-		\addplot [mark=,domain=-20:20,smooth]{e^(-( x^2 / 36.06737602222409 ))};
-		\addplot [mark=,domain=-5:5,smooth, gray,fill=gray!90!black,opacity=0.5]{e^(-( x^2 / 36.06737602222409 ))} \closedcycle;
+			ylabel=$Ranking$,
+			domain={-21:21},
+			samples at={-10,-9,...,-1,1,2,...,10}
+			]
+		\addplot [mark=x,smooth]{\normalPlotFWHM{x}{5}};
+		\addplot [mark=x,smooth,samples at={-10,-9,...,10}]{\normalPlotFWHMII{x}{3}};
 		\addplot [black,dotted] coordinates {(0,0) (0,1)};
 		\addplot [black,dashed] coordinates {(5,0) (5,1)};
+		\addplot [black,dashed] coordinates {(3,0) (3,1)};
 		\addplot [black,dashed] coordinates {(-5,0) (-5,1)};
+		\addplot [black,dashed] coordinates {(-3,0) (-3,1)};
 	\end{axis}
 \end{tikzpicture}
-\caption{Curve of the rating function $R_{d}$ with FWHM}
+%		\addplot [mark=,domain=-20:20,smooth]{e^(-(x^2/13.5))};
+%			x dir=reverse,
+\caption{Curves of the rating functions $R_{d}$ and $R_{s}$}
 \end{figure}
@@
 Unfortunately, the \emph{inverse document frequency} is not applicable to definition relations.
-In our approach, the frequency distribution of all terms, that participate in various definition relations is unknown beforehand.
+The frequency distribution of all terms, that participate in various relations is unknown.
 Thus, using tf-idf could affect the sensitivity of our rating statistic significantly.
 E.g., `Hamiltonian' as well as `length' are both valid entities.
-Obviously, the number of documents (respectively sentences) containing one of these words, within a corpus like Wikipedia, differs a lot.
+Obviously, the number of documents containing one of these words, within a corpus like Wikipedia, differs a lot.
 
-Therefore, we use the normalized term frequency. $t$ and $w$ are terms. $f$ denotes the raw frequency of a term within a document $d$.
+Therefore, we only use the normalized term frequency. $t$ and $w$ are terms. $f(t,d)$ denotes the frequency of a term within a document $d$.
 
@@
 
-We define our ranking statistic as the weighted mean of the rankings defined in Equations \ref{eq:dist}, \ref{eq:order}, and \ref{eq:tf}:
+We define our ranking statistic as the normalized weighted mean of the rankings defined in Equations \ref{eq:dist}, \ref{eq:order}, and \ref{eq:tf}:
 
@@
 	\label{eq:rating}
-	\frac{\alpha\cdot\mathrm{R}_{d}+\beta\cdot\mathrm{R}_{s}+\gamma\cdot\mathrm{tf}}{\alpha+\beta+\gamma} \mapsto [0;1]
+	R(n,\Delta,t,d)=\frac{\alpha{R}_{\mathrm d}(\Delta)+\beta{R}_{\mathrm s}(n)+\gamma\mathrm{tf}(t,d)}{\alpha+\beta+\gamma} \mapsto [0;1]
 \end{equation}
@@
 \subsection{Overview of the processor}
-The MLP processing system \cite{github} is a graph of PACTs, where data flows through.
+The MLP processing system \cite{github} is basically a set of PACTs, where data flows through.
 In contrast to other programming models like Map/Reduce, PACTs can operate on multiple inputs.
-At some point of time, we need an ordered list of tokenized sentences and a list of identified variables, to compute the ranking score.
+At some point of time, one needs an ordered list of tokenized sentences and a list of identified variables, to compute the ranking score.
 A \emph{CoGroup Contract} takes both lists and computes the subsets, build upon distinct document keys.
@@
 \begin{figure}[H]
-	\centering
 	\begin{tikzpicture}[node distance=5mm and 8mm]
@@
 		\node (Filter) [actor, right=of Candidates, align=center] {\emph{Reduce}\\\textbf{Filter}};
-		\node (Output) [right=of Filter,align=center]{Relation\ Candidates};
+		\node (Output) [right=of Filter,align=center]{Raw\ Candidates};
 		\draw[arrow] (Input)--(DocumentParser);
@@
 We used the \emph{WikiText} framework of the Mylyn project with a custom builder to transform the document into plain text.
-For analysing the \TeX \ mark-up in the \texttt{<math/>} tags, \emph{SnuggleTeX} transforms the formulas into their MathML representations, where variable names can be extracted easily.
-The second computation heavy part is tokenization and tagging of the plain text documents.
-We used the Stanford maximum entropy tagger with the \emph{WSJ left3words} language model.
+For analysing the \TeX \ mark-up in the \texttt{<math/>} tags, \emph{SnuggleTeX} transforms the formulas into their MathML representations, where variable names can be easily extracted.
+
+The second computation heavy part is tokenization and tagging of the plain text documents. We used the Stanford maximum entropy tagger with the \emph{WSJ left3words} language model.
+
 Finally, the rating for all nouns, contained in all candidate sentences, needs to be calculated.
@@
 First of all, people tend to incorrectly use \TeX\ trying to create formulas.
-I.e., \texttt{\textbackslash text\{log\}} is more often used than the correct operator \texttt{\textbackslash log}.
+E.g., \texttt{\textbackslash text\{log\}} is more often used than the correct operator \texttt{\textbackslash log}.
 Another problem is that sometimes people use indices as a form of annotation, like $T_{before}$, $R_{A}$, and $R_{B}$.
@@
 However, this approach is only applicable under certain conditions.
-In situations where both parts of a relation are unknown, other methods, especially supervised ones, are better used to bootstrap a language model.
-Unfortunately, we still need an other labelled test corpus to measure the real performance of a classifier, trained with the derived features.
+In situations where both parts of a relation are unknown, other methods, especially supervised ones, are used to bootstrap a language model.
+Unfortunately, we still need a labelled test corpus to measure the performance of a classifier, trained with our generated data.
 
 Apart from the mentioned weighting parameters, the ranking's quality is influenced by two further parameters.
-These are the assumptions about the parameters $c_{d}$ and $c_{s}$.
+These are the assumptions about the parameter $c$.
 Balancing all these parameters requires extensive knowledge about the underlying language.
-These parameters might also be specific for every type of relation.
+These parameters might be specific for every type of relation.
 Therefore, this strategy depends heavily on human input and many iterations of tuning.

### 2014-02-26 paper.tex
Little corrections of the abstract

@@
 In this paper, we present a solution for automated definition discovery for the variables occurring in mathematical formulae.
-Ordinary natural language processing approaches that are trained based on large manually annotated news corpora fall short in reasonable quality of definition discovery for scientific texts.
-In the \emph{Mathematical Language Processing} project we discover definition relations between variables and the surrounding text with a machine learning approach for classification.
-We extract the identifiers from the formulae and track co-occurrences of those in the surrounding text.
-The evaluation of our prototypical system, applied on the Wikipedia text corpus shows, that our approach augments the user experience for articles with mathematical formula substantially.
+Ordinary natural language processing approaches that are trained based on large, manually annotated news corpora fall short in reasonable quality of definition discovery for scientific texts.
+In the \emph{Mathematical Language Processing} project, we discover definitions %we did not discover any relations
+% a relation would be something like "let <subj> be a <object>", or "we call <subj> <obj>", or "<subject> is <object>" or just "<subj> <obj>"
+ for identifiers in formulae from the surrounding text by statistical methods.
+ %with a machine learning approach for classification. This was also not done in this paper
+%We extract the identifiers from the formulae and track co-occurrences of those in the surrounding text.
+The evaluation of our prototypical system, applied on the Wikipedia text corpus shows that our approach augments the user experience for articles with mathematical formula substantially.
 While hovering the identifiers occurring in the equation a pop-up with the most probable explanations for the identifiers occurs.
@@
 
-Unfortunately, no one has built a labelled text corpus that annotates variables and their definition.
-Instead of labelling large amounts of text by hand, we investigate a potential way of automatically labelling large corpora with a set of statistical assumptions.
+Unfortunately, no one has built a labeled text corpus that annotates variables and their definition.
+Instead of labeling large amounts of text by hand, we investigate a potential way of automatically labeling large corpora with a set of statistical assumptions.
 
 We chose the Wikipedia text corpus as a target because of two facts.
-First, the mark-up for editing formulas is a subset of \TeX.
+First, the mark-up for editing formulas is essentially a subset of \TeX.
 Therefore, it enables us to collect all variables within the set of formulas in a document.
@@
 
-Thereby we performed the normalisation, with regard to the model assumption that the probability to find a relation at $\Delta=1$ is maximal, i.e. $R(1)=1$. 
+Thereby we performed the normalization, with regard to the model assumption that the probability to find a relation at $\Delta=1$ is maximal, i.e. $R(1)=1$. 
 Manual evaluation of some articles indicated that $R(1)\approx\frac{1}{2}R(5)$.
@@
 Typically an author first introduces a term and uses it in subsequent sentences without necessarily repeating its definition.
-We assume that this kind of behaviour also applies to the use of variables.
+We assume that this kind of behavior also applies to the use of variables.
 Thus, we extend the idea of distance to the sequential order of sentences.
@@
 
+%460 identifiers
+
 \bibliography{mlp-papers}

### 2014-02-27 paper.tex
wrapped textblocks to be more readable in text editors

@@
 \begin{abstract}
-In this paper, we present a solution for automated definition discovery for the variables occurring in mathematical formulae.
-Ordinary natural language processing approaches that are trained based on large, manually annotated news corpora fall short in reasonable quality of definition discovery for scientific texts.
-In the \emph{Mathematical Language Processing} project, we discover definitions %we did not discover any relations
-% a relation would be something like "let <subj> be a <object>", or "we call <subj> <obj>", or "<subject> is <object>" or just "<subj> <obj>"
+In this paper, we present a solution for automated definition discovery for
+the variables occurring in mathematical formulae. Ordinary natural language
+processing approaches that are trained based on large, manually annotated news
+corpora fall short in reasonable quality of definition discovery for
+scientific texts. In the \emph{Mathematical Language Processing} project, we
+discover definitions
+%we did not discover any relations  a relation would be something like "let
+%<subj> be a <object>", or "we call <subj> <obj>", or "<subject> is <object>"
+%or just "<subj> <obj>"
 for identifiers in formulae from the surrounding text by statistical methods.
- %with a machine learning approach for classification. This was also not done in this paper
-%We extract the identifiers from the formulae and track co-occurrences of those in the surrounding text.
-The evaluation of our prototypical system, applied on the Wikipedia text corpus shows that our approach augments the user experience for articles with mathematical formula substantially.
-While hovering the identifiers occurring in the equation a pop-up with the most probable explanations for the identifiers occurs.
-The displayed definition information provides a good match to the actual meaning of the identifiers.
+%with a machine learning approach for classification. This was also not done
+%in this paper We extract the identifiers from the formulae and track co-
+%occurrences of those in the surrounding text.
+The evaluation of our prototypical system, applied on the Wikipedia text
+corpus shows that our approach augments the user experience for articles with
+mathematical formula substantially. While hovering the identifiers occurring
+in the equation a pop-up with the most probable explanations for the
+identifiers occurs. The displayed definition information provides a good match
+to the actual meaning of the identifiers.
 \keywords{machine learning, text mining, parallel computing}
@@
 \subsection{Motivation}
-Mathematical formulae are a viable source of information for a wide range of scientists.
-Often those formulas contain variables that are unknown or at least ambiguous to the reader.
-Therefore, one needs to study the surrounding text to find the relevant definition.
+Mathematical formulae are a viable source of information for a wide range of
+scientists. Often those formulas contain variables that are unknown or at
+least ambiguous to the reader. Therefore, one needs to study the surrounding
+text to find the relevant definition.
 
-An automatic information retrieval system can be used to reduce the readers effort to understand a formula. Especially students and scientists of other disciplines would profit from such a system.
+An automatic information retrieval system can be used to reduce the readers
+effort to understand a formula. Especially students and scientists of other
+disciplines would profit from such a system.
 
-Unfortunately, no one has built a labeled text corpus that annotates variables and their definition.
-Instead of labeling large amounts of text by hand, we investigate a potential way of automatically labeling large corpora with a set of statistical assumptions.
+Unfortunately, no one has built a labeled text corpus that annotates variables
+and their definition. Instead of labeling large amounts of text by hand, we
+investigate a potential way of automatically labeling large corpora with a set
+of statistical assumptions.
 
-We chose the Wikipedia text corpus as a target because of two facts.
-First, the mark-up for editing formulas is essentially a subset of \TeX.
-Therefore, it enables us to collect all variables within the set of formulas in a document.
-Second, we can leverage the semantic information between mark-up and word tokens, e.g., hyperlinks.
+We chose the Wikipedia text corpus as a target because of two facts. First,
+the mark-up for editing formulas is essentially a subset of \TeX. Therefore,
+it enables us to collect all variables within the set of formulas in a
+document. Second, we can leverage the semantic information between mark-up and
+word tokens, e.g., hyperlinks.
 
-The English Wikipedia contains roughly four million articles.
-Even if we only pick articles containing \texttt{<math/>} mark-up, our processor still needs to compute tens of thousands of articles.
-Especially when using a maximum entropy \emph{POS} tagger \cite{Rathna96}, like the one in Stanford's NLP framework, one can make use of a parallel processing system to speed up computation.
+The English Wikipedia contains roughly four million articles. Even if we only
+pick articles containing \texttt{<math/>} mark-up, our processor still needs
+to compute tens of thousands of articles. Especially when using a maximum
+entropy \emph{POS} tagger \cite{Rathna96}, like the one in Stanford's NLP
+framework, one can make use of a parallel processing system to speed up
+computation.
 
-We will show how the proposed strategy can be implemented in the PACT programming model \cite{Alexandrov2010}.
+We will show how the proposed strategy can be implemented in the PACT
+programming model \cite{Alexandrov2010}.
 
 \subsection{Related Work}
-\citeauthor{Quoc2010} (2010) proposed in \cite{Quoc2010} an approach for relating whole formulas to sentences and their describing paragraphs.
-\citeauthor{Yokoi} (2010) trained a \emph{support vector machine} to extract natural language descriptions for mathematical expressions \cite{Yokoi}.
+\citeauthor{Quoc2010} (2010) proposed in cite{Quoc2010} an approach for
+relating whole formulas to sentences and their describing paragraphs.
+\citeauthor{Yokoi} (2010) trained a emph{support vector machine} to extract
+natural language descriptions for mathematical expressions cite{Yokoi}.
 
 \section{Relation Detection}
-The basic assumption of our approach is that if two entities take part in a relation, those two co-occur in the same sentence.
-This assumption is also used in the \emph{Distant Supervision} \cite{Mintz2008} approach by \citeauthor{Mintz2008}.
-Furthermore, we assume in a scientific document, which contains a specific variable name, there is at least one sentence with such a relation.
+The basic assumption of our approach is that if two entities take part in a
+relation, those two co-occur in the same sentence. This assumption is also
+used in the \emph{Distant Supervision} \cite{Mintz2008} approach by
+\citeauthor{Mintz2008}. Furthermore, we assume in a scientific document, which
+contains a specific variable name, there is at least one sentence with such a
+relation.
 
 The two entities have some characteristics that can be used for filtering.
-First of all, the relevant variables can be retrieved accurately from \TeX\ formatted formulas (see Section \ref{vr}).
-The fact that a relation consists of two entities and we know that one must be the variable, leads to an obvious conclusion: Sentences without a variable cannot contain a definition relation and can be discarded.
-Second, one can assume that the other entity must be a noun or a compound noun. Thus, all words that are not nouns or adjectives describing these nouns, can also be discarded.
+First of all, the relevant variables can be reieved accurately from \TeX\
+formatted formulas (see Section \ref{vr}). The fact that a relation consists
+of two entities and we know that one must be the variable, leads to an obvious
+conclusion: Sentences without a variable cannot contain a definition relation
+and can be discarded. Second, one can assume that the other entity must be a
+noun or a compound noun. Thus, all words that are not nouns or adjectives
+describing these nouns, can also be discarded.
 
 \subsection{Numerical Statistics}
-As a prerequisite for detecting potential relations, one needs to identify all occurrences of a variable within a document.
-We assume that the distance between the identifier and the corresponding definition is normal distributed.
+As a prerequisite for detecting potential relations, one needs to identify all
+occurrences of a variable within a document. We assume that the distance
+between the identifier and the corresponding definition is normal distributed.
 
-Formally, we calculate the distance in words $\Delta=p_\mathrm{def}-p_\mathrm{id}$  between $p_\mathrm{def}$ the position of the definition and $p_\mathrm{id}$ the position of the identifier.
+Formally, we calculate the distance in words
+$\Delta=p_\mathrm{def}-p_\mathrm{id}$  between $p_\mathrm{def}$ the position
+of the definition and $p_\mathrm{id}$ the position of the identifier.
 
@@
 
-Thereby we performed the normalization, with regard to the model assumption that the probability to find a relation at $\Delta=1$ is maximal, i.e. $R(1)=1$. 
-Manual evaluation of some articles indicated that $R(1)\approx\frac{1}{2}R(5)$.
-As, the full width half maximum is at $\Delta=\sqrt{2 \ln(2)  \sigma^2+1}$, and the standard deviation evaluates to $\sigma(\Delta)=\sqrt\frac{\Delta^2-1}{2\ln 2}$.
-As a consequence we choose $\sigma_d=\sqrt\frac{5^2-1}{2\ln 2}$ and write $R_\mathrm{d}(\Delta) \equiv R_{\sigma_\mathrm{d}}(\Delta) $.
-Please note that this assumption is heavily influenced by the language of the text corpus.
+Thereby we performed the normalization, with regard to the model assumption
+that the probability to find a relation at $\Delta=1$ is maximal, i.e.
+$R(1)=1$.  Manual evaluation of some articles indicated that
+$R(1)\approx\frac{1}{2}R(5)$. As, the full width half maximum is at
+$\Delta=\sqrt{2 \ln(2)  \sigma^2+1}$, and the standard deviation evaluates to
+$\sigma(\Delta)=\sqrt\frac{\Delta^2-1}{2\ln 2}$. As a consequence we choose
+$\sigma_d=\sqrt\frac{5^2-1}{2\ln 2}$ and write $R_\mathrm{d}(\Delta) \equiv
+R_{\sigma_\mathrm{d}}(\Delta) $. Please note that this assumption is heavily
+influenced by the language of the text corpus.
 
-Not only the distances within one sentence are of interest.
-Also the sequential order of sentences, containing a variable, carry some viable sources for statistics.
-Typically an author first introduces a term and uses it in subsequent sentences without necessarily repeating its definition.
-We assume that this kind of behavior also applies to the use of variables.
-Thus, we extend the idea of distance to the sequential order of sentences.
-Here, the distance is the count $n$ of candidate sentences, beginning at the top of the document.
+Not only the distances within one sentence are of interest. Also the
+sequential order of sentences, containing a variable, carry some viable
+sources for statistics. Typically an author first introduces a term and uses
+it in subsequent sentences without necessarily repeating its definition. We
+assume that this kind of behavior also applies to the use of variables. Thus,
+we extend the idea of distance to the sequential order of sentences. Here, the
+distance is the count $n$ of candidate sentences, beginning at the top of the
+document.
 
-In our experiments, we expect that half of the definition relations are found in the sentence with 3 sentences difference. Thus we choose 
+In our experiments, we expect that half of the definition relations are found
+in the sentence with 3 sentences difference. Thus we choose
 \begin{equation}
@@
 
-Furthermore, we take distributional properties into account.
-The classic tf-idf \cite{Salton86} statistic reflects the importance of a term to a document.
-Unfortunately, the \emph{inverse document frequency} is not applicable to definition relations.
-The frequency distribution of all terms, that participate in various relations is unknown.
-Thus, using tf-idf could affect the sensitivity of our rating statistic significantly.
-E.g., `Hamiltonian' as well as `length' are both valid entities.
-Obviously, the number of documents containing one of these words, within a corpus like Wikipedia, differs a lot.
+Furthermore, we take distributional properties into account. The classic tf-
+idf \cite{Salton86} statistic reflects the importance of a term to a document.
+Unfortunately, the \emph{inverse document frequency} is not applicable to
+definition relations. The frequency distribution of all terms, that
+participate in various relations is unknown. Thus, using tf-idf could affect
+the sensitivity of our rating statistic significantly. E.g., `Hamiltonian' as
+well as `length' are both valid entities. Obviously, the number of documents
+containing one of these words, within a corpus like Wikipedia, differs a lot.
 
-Therefore, we only use the normalized term frequency. $t$ and $w$ are terms. $f(t,d)$ denotes the frequency of a term within a document $d$.
+Therefore, we only use the normalized term frequency. $t$ and $w$ are terms.
+$f(t,d)$ denotes the frequency of a term within a document $d$.
 
@@
 
-We define our ranking statistic as the normalized weighted mean of the rankings defined in Equations \ref{eq:dist}, \ref{eq:order}, and \ref{eq:tf}:
+We define our ranking statistic as the normalized weighted mean of the
+rankings defined in Equations \ref{eq:dist}, \ref{eq:order}, and \ref{eq:tf}:
 
@@
 \subsection{Overview of the processor}
-The MLP processing system \cite{github} is basically a set of PACTs, where data flows through.
-In contrast to other programming models like Map/Reduce, PACTs can operate on multiple inputs.
-At some point of time, one needs an ordered list of tokenized sentences and a list of identified variables, to compute the ranking score.
-A \emph{CoGroup Contract} takes both lists and computes the subsets, build upon distinct document keys.
-Finally, a \emph{Reduce Contract} filters out results below a given threshold $\theta$.
+The MLP processing system \cite{github} is basically a set of PACTs, where
+data flows through. In contrast to other programming models like Map/Reduce,
+PACTs can operate on multiple inputs. At some point of time, one needs an
+ordered list of tokenized sentences and a list of identified variables, to
+compute the ranking score. A \emph{CoGroup Contract} takes both lists and
+computes the subsets, build upon distinct document keys. Finally, a
+\emph{Reduce Contract} filters out results below a given threshold $\theta$.
 
@@
 \end{figure}
-There are three computation heavy parts in the processor.
-First, the raw Wikipedia document dumps need to be transformed into plain text.
-At the same step, the formulas need to be evaluated and replaced by place-holder words.
-We used the \emph{WikiText} framework of the Mylyn project with a custom builder to transform the document into plain text.
-For analysing the \TeX \ mark-up in the \texttt{<math/>} tags, \emph{SnuggleTeX} transforms the formulas into their MathML representations, where variable names can be easily extracted.
 
-The second computation heavy part is tokenization and tagging of the plain text documents. We used the Stanford maximum entropy tagger with the \emph{WSJ left3words} language model.
+There are three computation heavy parts in the processor. First, the raw
+Wikipedia document dumps need to be transformed into plain text. At the same
+step, the formulas need to be evaluated and replaced by place-holder words. We
+used the \emph{WikiText} framework of the Mylyn project with a custom builder
+to transform the document into plain text. For analysing the \TeX \ mark-up in
+the \texttt{<math/>} tags, \emph{SnuggleTeX} transforms the formulas into
+their MathML representations, where variable names can be easily extracted.
 
-Finally, the rating for all nouns, contained in all candidate sentences, needs to be calculated.
+The second computation heavy part is tokenization and tagging of the plain
+text documents. We used the Stanford maximum entropy tagger with the \emph{WSJ
+left3words} language model.
+
+Finally, the rating for all nouns, contained in all candidate sentences, needs
+to be calculated.
 
@@
 \label{vr}
-Throughout our experiments we made some observations that had an impact on the accuracy of retrieving the correct set of variables.
-First of all, people tend to incorrectly use \TeX\ trying to create formulas.
-E.g., \texttt{\textbackslash text\{log\}} is more often used than the correct operator \texttt{\textbackslash log}.
-Another problem is that sometimes people use indices as a form of annotation, like $T_{before}$, $R_{A}$, and $R_{B}$.
+Throughout our experiments we made some observations that had an impact on the
+accuracy of retrieving the correct set of variables. First of all, people tend
+to incorrectly use \TeX\ trying to create formulas. E.g.,
+\texttt{\textbackslash text\{log\}} is more often used than the correct
+operator \texttt{\textbackslash log}. Another problem is that sometimes people
+use indices as a form of annotation, like $T_{before}$, $R_{A}$, and $R_{B}$.
 These ambiguities lead to a increased number of false positives.
@@
 We took a very conservative approach and preprocessed all formulas.
-\texttt{\textbackslash text\{\}} blocks along with plain subscriptions will be removed before analysis.
-Moreover, we created a comprehensive blacklist to improve the results further.
-Short words like `a' and `i', which are very common, could be easily matched by our processor if a formula contains one of those symbols.
-Additionally, we blacklist common mathematical operators, constants, and functions.
+\texttt{\textbackslash text\{\}} blocks along with plain subscriptions will be
+removed before analysis. Moreover, we created a comprehensive blacklist to
+improve the results further. Short words like `a' and `i', which are very
+common, could be easily matched by our processor if a formula contains one of
+those symbols. Additionally, we blacklist common mathematical operators,
+constants, and functions.
 
-Because of the fact, that we do not have any annotated test corpora, evaluation has to be performed by hand.
-Therefore, we took a sample of 30 random documents and counted all matches.
-The resulting estimates for \emph{recall} and \emph{precision} are 0.99 and 0.86 respectively.
+Because of the fact, that we do not have any annotated test corpora,
+evaluation has to be performed by hand. Therefore, we took a sample of 30
+random documents and counted all matches. The resulting estimates for
+\emph{recall} and \emph{precision} are 0.99 and 0.86 respectively.
 
 \subsubsection{Tuning}
-The weighting parameters $\alpha$, $\beta$, and $\gamma$ in Equation \ref{eq:rating} are used to balance the contribution of all embedded statistics to the overall ranking.
-But these parameters need to be tuned carefully.
-Otherwise the results tend to be biased towards one parameter.
-In other words, if the distance measure $R_{D}$ dominates the other measures, more complex\footnote{in terms of words between both entities, participating in a relation} relations are less likely to be discovered.
+The weighting parameters $\alpha$, $\beta$, and $\gamma$ in Equation
+\ref{eq:rating} are used to balance the contribution of all embedded
+statistics to the overall ranking. But these parameters need to be tuned
+carefully. Otherwise the results tend to be biased towards one parameter. In
+other words, if the distance measure $R_{D}$ dominates the other measures,
+more complex\footnote{in terms of words between both entities, participating
+in a relation} relations are less likely to be discovered.
 
-The \emph{term frequency} turns out to be the least robust statistic.
-Usually, the frequency of a specific term co-occurring with a variable name within one document is rather low.
-Thus, the derived ranking is relatively volatile.
-In our experiments, we chose $\gamma$ to be 0.75 ($\alpha = \beta = 1$) for a more robust ranking.
+The \emph{term frequency} turns out to be the least robust statistic. Usually,
+the frequency of a specific term co-occurring with a variable name within one
+document is rather low. Thus, the derived ranking is relatively volatile. In
+our experiments, we chose $\gamma$ to be 0.75 ($\alpha = \beta = 1$) for a
+more robust ranking.
 
 \section{Conclusion}
-Our experiments showed that combining a POS tagger with numerical statistics about the text surface, can lead to quality results.
-However, this approach is only applicable under certain conditions.
-In situations where both parts of a relation are unknown, other methods, especially supervised ones, are used to bootstrap a language model.
-Unfortunately, we still need a labelled test corpus to measure the performance of a classifier, trained with our generated data.
+Our experiments showed that combining a POS tagger with numerical statistics
+about the text surface, can lead to quality results. However, this approach is
+only applicable under certain conditions. In situations where both parts of a
+relation are unknown, other methods, especially supervised ones, are used to
+bootstrap a language model. Unfortunately, we still need a labelled test
+corpus to measure the performance of a classifier, trained with our generated
+data.
 
-Apart from the mentioned weighting parameters, the ranking's quality is influenced by two further parameters.
-These are the assumptions about the parameter $c$.
-Balancing all these parameters requires extensive knowledge about the underlying language.
-These parameters might be specific for every type of relation.
-Therefore, this strategy depends heavily on human input and many iterations of tuning.
+Apart from the mentioned weighting parameters, the ranking's quality is
+influenced by two further parameters. These are the assumptions about the
+parameter $c$. Balancing all these parameters requires extensive knowledge
+about the underlying language. These parameters might be specific for every
+type of relation. Therefore, this strategy depends heavily on human input and
+many iterations of tuning.
 
-To further improve the robustness of the \emph{term frequency} measure, one might cluster sets of scientific documents, based on their specific field of research.
+To further improve the robustness of the \emph{term frequency} measure, one
+might cluster sets of scientific documents, based on their specific field of
+research.
 

### 2014-03-10 paper.tex
Shrink down the paper to fit to four pages

@@
 \begin{abstract}
-In this paper, we present a solution for automated definition discovery for
-the variables occurring in mathematical formulae. Ordinary natural language
-processing approaches that are trained based on large, manually annotated news
-corpora fall short in reasonable quality of definition discovery for
+This paper presents automated definition discovery for
+variables occurring in mathematical formulae. Ordinary natural language
+processing approaches are trained based on large, manually annotated news
+corpora. They fall short in reasonable quality of definition discovery for
 scientific texts. In the \emph{Mathematical Language Processing} project, we
@@
 
-\subsection{Motivation}
 Mathematical formulae are a viable source of information for a wide range of
@@
 
-We will show how the proposed strategy can be implemented in the PACT
+We implement  show how the proposed strategy can be implemented in the PACT
 programming model \cite{Alexandrov2010}.
 
-\subsection{Related Work}
-\citeauthor{Quoc2010} (2010) proposed in cite{Quoc2010} an approach for
+\paragraph{Related Work}
+\citeauthor{Quoc2010}\cite{Quoc2010} proposed an approach for
 relating whole formulas to sentences and their describing paragraphs.
-\citeauthor{Yokoi} (2010) trained a emph{support vector machine} to extract
-natural language descriptions for mathematical expressions cite{Yokoi}.
+\citeauthor{Yokoi}\cite{Yokoi} trained a emph{support vector machine} to extract
+natural language descriptions for mathematical expressions.
 
-\section{Relation Detection}
-The basic assumption of our approach is that if two entities take part in a
-relation, those two co-occur in the same sentence. This assumption is also
-used in the \emph{Distant Supervision} \cite{Mintz2008} approach by
-\citeauthor{Mintz2008}. Furthermore, we assume in a scientific document, which
-contains a specific variable name, there is at least one sentence with such a
-relation.
+\section{Statistical definition discovery}
+We detect relations between identifiers and their description in two steps.
+First, we extract the identifiers from the display-style formulae and
+second we determine their description from the surrounding text.
 
-The two entities have some characteristics that can be used for filtering.
-First of all, the relevant variables can be reieved accurately from \TeX\
-formatted formulas (see Section \ref{vr}). The fact that a relation consists
-of two entities and we know that one must be the variable, leads to an obvious
-conclusion: Sentences without a variable cannot contain a definition relation
-and can be discarded. Second, one can assume that the other entity must be a
-noun or a compound noun. Thus, all words that are not nouns or adjectives
-describing these nouns, can also be discarded.
+Inspired by the \emph{Distant Supervision} approach \cite{Mintz2008},
+we assume that the identifier and the definition co-occur in the same sentence.
+Furthermore, we restrict ourselves to identifiers that co-occur with a noun
+phrase in one sentence in the document at least.
 
-\subsection{Numerical Statistics}
-As a prerequisite for detecting potential relations, one needs to identify all
-occurrences of a variable within a document. We assume that the distance
-between the identifier and the corresponding definition is normal distributed.
+For each description candidate we use the weighted sum
+\begin{equation} \label{eq:rating}
+	R(n,\Delta,t,d)=\frac{\alpha{R}_{\sigma_\mathrm d}(\Delta)
+		+\beta{R}_{\sigma_\mathrm s}(n)
+		+\gamma\mathrm{tf}(t,d)}{\alpha+\beta+\gamma} \mapsto [0;1]
+\end{equation}
+that respects the distance $\Delta$ between identifier and definition term $t$, the distance $n$ between the display-style equation and th sentence that contains the term and the identifier and the term frequency $\mathrm{tf}(t,d)$ in the current document $d$
+with a normal distributed distance measure $R_\sigma(\Delta)= \exp\left[-\frac{1}{2}\frac{\Delta^2-1}{\sigma^2}\right].$
+Our model assumption is that the probability to find a relation at $\Delta=1$ like for example in \emph{the energy $E$ and the mass $m$} is maximal, i.e.
+$R_{\sigma_\mathrm d}(1)=1$.
+In order to determine the full width half maximum of our distribution we evaluated some articles manually and found $R_{\sigma_\mathrm d}(1)\approx 2 R_{\sigma_\mathrm d}(5)$ and thus $\sigma_d=\sqrt\frac{5^2-1}{2\ln 2}$.
 
-Formally, we calculate the distance in words
-$\Delta=p_\mathrm{def}-p_\mathrm{id}$  between $p_\mathrm{def}$ the position
-of the definition and $p_\mathrm{id}$ the position of the identifier.
+The probability to find a correct definition decays for 50 \% within three sentences. Consequently  $\sigma_\mathrm s=2\left({\ln 2}\right)^{-\frac{1}{2}}$.
 
-Thus, the assumed normalized relation probability density function reads
-\begin{equation}
-\label{eq:dist}
-R_\sigma(\Delta)= {\sigma \sqrt{2\pi}}{e^{{\left(2\sigma^2\right)^{-1}}}} \mathcal N(0,\sigma^2)(\Delta)\equiv \exp\left[-\frac{1}{2}\frac{\Delta^2-1}{\sigma^2}\right].
-\end{equation}
-\newcommand{\normalPlot}[2]{(e^((-1/2)*(((#1)^2-1)/(#2)^2)))} %1/((#2)*sqrt(2*pi)))*
-\newcommand{\normalPlotFWHM}[2]{\normalPlot{#1}{ sqrt(((#2)^2-1)/(2*ln(2)))}}
+The classic tf-idf \cite{Salton86} statistic reflects the importance of a term to a document.
+For our task the inverse document frequency (idf)  would assign high penalties to frequent which are valid definitions for identifiers like in the example `Energy` and `Einstein`.
+Consequently we used just the term frequency which still had an impact on the overall result that seemed to high in our manual experiments.
+Therefore, we use the tuning parameters $\gamma=0.75$ and $\alpha = \beta = 1$.
 
-Thereby we performed the normalization, with regard to the model assumption
-that the probability to find a relation at $\Delta=1$ is maximal, i.e.
-$R(1)=1$.  Manual evaluation of some articles indicated that
-$R(1)\approx\frac{1}{2}R(5)$. As, the full width half maximum is at
-$\Delta=\sqrt{2 \ln(2)  \sigma^2+1}$, and the standard deviation evaluates to
-$\sigma(\Delta)=\sqrt\frac{\Delta^2-1}{2\ln 2}$. As a consequence we choose
-$\sigma_d=\sqrt\frac{5^2-1}{2\ln 2}$ and write $R_\mathrm{d}(\Delta) \equiv
-R_{\sigma_\mathrm{d}}(\Delta) $. Please note that this assumption is heavily
-influenced by the language of the text corpus.
-
-Not only the distances within one sentence are of interest. Also the
-sequential order of sentences, containing a variable, carry some viable
-sources for statistics. Typically an author first introduces a term and uses
-it in subsequent sentences without necessarily repeating its definition. We
-assume that this kind of behavior also applies to the use of variables. Thus,
-we extend the idea of distance to the sequential order of sentences. Here, the
-distance is the count $n$ of candidate sentences, beginning at the top of the
-document.
-
-In our experiments, we expect that half of the definition relations are found
-in the sentence with 3 sentences difference. Thus we choose
-\begin{equation}
-\label{eq:order}
-R_{s}(n)={\sigma_\mathrm{s} \sqrt{2\pi}} \mathcal N(0,\sigma_\mathrm{s}^2)(n)= \exp\left[-\frac{1}{2}\frac{n^2}{\sigma_\mathrm{s}^2}\right]
-\end{equation}
- with $\sigma_\mathrm{s}=\frac{3}{\sqrt{2\ln 2}}$.
-\newcommand{\normalPlotFWHMII}[2]{(e^((-1/2)*(((#1)^2*2*ln(2))/(#2)^2)))}
-
-\begin{figure}[H]
-\begin{tikzpicture}
-	\begin{axis}[
-			axis x line = bottom,
-			axis y line = left,
-			height=3.5cm, width=12cm,
-			xmin=-10, xmax=10,
-			ymin=0, ymax=1,
-			xtick={-5,-3,0,3,5},
-			ytick={0,0.5,1},
-			xlabel=$Distance$,
-			ylabel=$Ranking$,
-			domain={-21:21},
-			samples at={-10,-9,...,-1,1,2,...,10}
-			]
-		\addplot [mark=x,smooth]{\normalPlotFWHM{x}{5}};
-		\addplot [mark=x,smooth,samples at={-10,-9,...,10}]{\normalPlotFWHMII{x}{3}};
-		\addplot [black,dotted] coordinates {(0,0) (0,1)};
-		\addplot [black,dashed] coordinates {(5,0) (5,1)};
-		\addplot [black,dashed] coordinates {(3,0) (3,1)};
-		\addplot [black,dashed] coordinates {(-5,0) (-5,1)};
-		\addplot [black,dashed] coordinates {(-3,0) (-3,1)};
-	\end{axis}
-\end{tikzpicture}
-%		\addplot [mark=,domain=-20:20,smooth]{e^(-(x^2/13.5))};
-%			x dir=reverse,
-\caption{Curves of the rating functions $R_{d}$ and $R_{s}$}
-\end{figure}
-
-Furthermore, we take distributional properties into account. The classic tf-
-idf \cite{Salton86} statistic reflects the importance of a term to a document.
-Unfortunately, the \emph{inverse document frequency} is not applicable to
-definition relations. The frequency distribution of all terms, that
-participate in various relations is unknown. Thus, using tf-idf could affect
-the sensitivity of our rating statistic significantly. E.g., `Hamiltonian' as
-well as `length' are both valid entities. Obviously, the number of documents
-containing one of these words, within a corpus like Wikipedia, differs a lot.
-
-Therefore, we only use the normalized term frequency. $t$ and $w$ are terms.
-$f(t,d)$ denotes the frequency of a term within a document $d$.
-
-\begin{equation}
-	\label{eq:tf}
-	\mathrm{tf}(t,d) = \frac{\mathrm{f}(t,d)}{\max\{\mathrm{f}(w,d):w \in d\}}
-\end{equation}
-
-We define our ranking statistic as the normalized weighted mean of the
-rankings defined in Equations \ref{eq:dist}, \ref{eq:order}, and \ref{eq:tf}:
-
-\begin{equation}
-	\label{eq:rating}
-	R(n,\Delta,t,d)=\frac{\alpha{R}_{\mathrm d}(\Delta)+\beta{R}_{\mathrm s}(n)+\gamma\mathrm{tf}(t,d)}{\alpha+\beta+\gamma} \mapsto [0;1]
-\end{equation}
+%The weighting parameters $\alpha$, $\beta$, and $\gamma$ in Equation
+%\ref{eq:rating} are used to balance the contribution of all embedded
+%statistics to the overall ranking. But these parameters need to be tuned
+%carefully. Otherwise the results tend to be biased towards one parameter. In
+%other words, if the distance measure $R_{D}$ dominates the other measures,
+%more complex\footnote{in terms of words between both entities, participating
+%in a relation} relations are less likely to be discovered.
+%
+%The \emph{term frequency} turns out to be the least robust statistic. Usually,
+%the frequency of a specific term co-occurring with a variable name within one
+%document is rather low. Thus, the derived ranking is relatively volatile. In
+%our experiments, we chose $\gamma$ to be 0.75 ($\alpha = \beta = 1$) for a
+%more robust ranking.
+%to  E.g., `Hamiltonian' as well as `length' are both valid entities. Obviously, the number of documents containing one of these words, within a corpus like Wikipedia, differs a lot.
+%
+%Therefore, we only use the normalized term frequency. $t$ and $w$ are terms.
+%$f(t,d)$ denotes the frequency of a term within a document $d$.
+%
+%\begin{equation}
+%	\label{eq:tf}
+%	\mathrm{tf}(t,d) = \frac{\mathrm{f}(t,d)}{\max\{\mathrm{f}(w,d):w \in d\}}
+%\end{equation}
+%
 
 \section{Implementation}
-
-\subsection{Overview of the processor}
 The MLP processing system \cite{github} is basically a set of PACTs, where
@@
 
-\subsection{Experiments}
-
-\subsubsection{Variable Retrieval}
+%\subsection{Experiments}
+\section{Evaluation}
+%\subsubsection{Variable Retrieval}
 \label{vr}
@@
 
-\subsubsection{Tuning}
-The weighting parameters $\alpha$, $\beta$, and $\gamma$ in Equation
-\ref{eq:rating} are used to balance the contribution of all embedded
-statistics to the overall ranking. But these parameters need to be tuned
-carefully. Otherwise the results tend to be biased towards one parameter. In
-other words, if the distance measure $R_{D}$ dominates the other measures,
-more complex\footnote{in terms of words between both entities, participating
-in a relation} relations are less likely to be discovered.
-
-The \emph{term frequency} turns out to be the least robust statistic. Usually,
-the frequency of a specific term co-occurring with a variable name within one
-document is rather low. Thus, the derived ranking is relatively volatile. In
-our experiments, we chose $\gamma$ to be 0.75 ($\alpha = \beta = 1$) for a
-more robust ranking.
-
 \section{Conclusion}
@@
 %460 identifiers
+\begingroup
+\let\clearpage\relax
+\bibliography{mlp-papers}
+\endgroup
 
-\bibliography{mlp-papers}
 \end{document}

### 2014-03-12 paper.tex
first complete draft of short paper

@@
 \usepackage{float}
+\usepackage{todonotes}
 
@@
 \usepackage{url}
-\urldef{\mailsa}\path|rob@clabs.cc, schubotz@tu-berlin.de|
+\urldef{\mailsa}\path|rob@clabs.cc,schubotz@tu-berlin.de |
 \newcommand{\todaydate}{\leadingzero{\day}.\leadingzero{\month}.\the\year}
@@
 \begin{abstract}
-This paper presents automated definition discovery for
-variables occurring in mathematical formulae. Ordinary natural language
-processing approaches are trained based on large, manually annotated news
-corpora. They fall short in reasonable quality of definition discovery for
-scientific texts. In the \emph{Mathematical Language Processing} project, we
-discover definitions
-%we did not discover any relations  a relation would be something like "let
-%<subj> be a <object>", or "we call <subj> <obj>", or "<subject> is <object>"
-%or just "<subj> <obj>"
-for identifiers in formulae from the surrounding text by statistical methods.
-%with a machine learning approach for classification. This was also not done
-%in this paper We extract the identifiers from the formulae and track co-
-%occurrences of those in the surrounding text.
+In natural language words and phrases themselves imply the semantics.
+Complimentary, the meaning of identifiers in mathematical formulae is undefined.
+Thus scientist must study the context to decode the meaning.
+The \emph{Mathematical Language Processing} project, aims to support that process and
+presents automated identifier definition discovery.
+
+Ordinary natural language processing approaches are trained on large, manually annotated news
+corpora. They fall short in reasonable quality of definition discovery for scientific texts.
+We use statistical methods to discover definitions in the surrounding text.
+
 The evaluation of our prototypical system, applied on the Wikipedia text
-corpus shows that our approach augments the user experience for articles with
-mathematical formula substantially. While hovering the identifiers occurring
-in the equation a pop-up with the most probable explanations for the
-identifiers occurs. The displayed definition information provides a good match
-to the actual meaning of the identifiers.
+corpus shows that our approach augments the user experience
+substantially. While hovering the identifiers occurring
+in the formula tool-tips with the most probable definitions
+occur.
+Tests with random samples show, that the displayed definitions provide
+a good match with the actual meaning of the identifiers.   
 \keywords{machine learning, text mining, parallel computing}
@@
 Mathematical formulae are a viable source of information for a wide range of
-scientists. Often those formulas contain variables that are unknown or at
-least ambiguous to the reader. Therefore, one needs to study the surrounding
-text to find the relevant definition.
+scientists. Often, they contain variables that are unknown or ambiguous to the reader.
+Therefore, one needs to study the surrounding text to find the relevant definition.
 
 An automatic information retrieval system can be used to reduce the readers
-effort to understand a formula. Especially students and scientists of other
-disciplines would profit from such a system.
+effort. Especially students and scientists of other
+disciplines would profit from a system that helps them to understand formulae.
 
-Unfortunately, no one has built a labeled text corpus that annotates variables
-and their definition. Instead of labeling large amounts of text by hand, we
-investigate a potential way of automatically labeling large corpora with a set
-of statistical assumptions.
+To build such a system a labeled text corpus that annotates variables
+and their definition is desirable. At the project start, such a corpus was not available.
+%Unfortunately, no one has built . Instead of labeling large amounts of text by hand,
+Consequently, we use simple statistical assumptions about co-occurrences between identifiers
+and their potential definition. In combination with the Stratosphere system
+that is capable of scalable data processing,  we are able to generate a large amount of 
+definition relation candidates in a reasonable and a little implementation effort.
+% the data processing system Stratosphere we are able to analyze large text corpora   and a system that  investigate a potential way of automatically labeling large corpora with a set
+%of statistical assumptions.
 
-We chose the Wikipedia text corpus as a target because of two facts. First,
-the mark-up for editing formulas is essentially a subset of \TeX. Therefore,
-it enables us to collect all variables within the set of formulas in a
-document. Second, we can leverage the semantic information between mark-up and
+We chose the Wikipedia text corpus as a target because of two facts.
+First, it is easy to extract the formulas and identifiers from the text,
+since the input language for formula texvc is very restrictive.
+Second, we can leverage the semantic information between mark-up and
 word tokens, e.g., hyperlinks.
@@
 The English Wikipedia contains roughly four million articles. Even if we only
-pick articles containing \texttt{<math/>} mark-up, our processor still needs
+pick articles containing \texttt{<math/>} tag, our processor still needs
 to compute tens of thousands of articles. Especially when using a maximum
@@
 
-We implement  show how the proposed strategy can be implemented in the PACT
-programming model \cite{Alexandrov2010}.
+Therefore we implement the proposed strategy with the Stratosphere system based
+on the PACT programming model \cite{Alexandrov2010}.
 
@@
 relating whole formulas to sentences and their describing paragraphs.
-\citeauthor{Yokoi}\cite{Yokoi} trained a emph{support vector machine} to extract
+\citeauthor{Yokoi}\cite{Yokoi} trained a \emph{support vector machine} to extract
 natural language descriptions for mathematical expressions.
@@
 
-\section{Implementation}
-The MLP processing system \cite{github} is basically a set of PACTs, where
-data flows through. In contrast to other programming models like Map/Reduce,
-PACTs can operate on multiple inputs. At some point of time, one needs an
-ordered list of tokenized sentences and a list of identified variables, to
-compute the ranking score. A \emph{CoGroup Contract} takes both lists and
-computes the subsets, build upon distinct document keys. Finally, a
-\emph{Reduce Contract} filters out results below a given threshold $\theta$.
-
-\tikzset{actor/.style={
-        rectangle,
-        minimum size=6mm,
-        very thick,
-        draw=gray!50!black!50,
-        top color=white,
-        bottom color=gray!50!black!20
-    },
-    arrow/.style={
-        -latex, thick, shorten <=2pt,shorten >=2pt
-    }
-}
-\begin{figure}[H]
-	\begin{tikzpicture}[node distance=5mm and 8mm]
-		\node (Input) [align=center]{Wiki Dumps};
-		\node (DocumentParser) [actor, right=of Input, align=center] {\emph{Map}\\\textbf{Parser}};
-		\node (Candidates) [actor, right=of DocumentParser, align=center] {\emph{CoGroup}\\\textbf{Kernel}};
-		\node (Sentence) [actor, above=of DocumentParser, align=center] {\emph{Map}\\\textbf{Tagger}};
-		\node (Filter) [actor, right=of Candidates, align=center] {\emph{Reduce}\\\textbf{Filter}};
-		\node (Output) [right=of Filter,align=center]{Raw\ Candidates};
-		\draw[arrow] (Input)--(DocumentParser);
-		\draw[arrow] (DocumentParser)--(Sentence);
-		\draw[arrow] (DocumentParser)--(Candidates);
-		\draw[arrow] (Sentence.east)--(Candidates.north);
-		\draw[arrow] (Candidates)--(Filter);
-		\draw[arrow] (Filter)--(Output);
-	\end{tikzpicture}
-\caption{Data flow of the PACT program}
-\end{figure}
-
-There are three computation heavy parts in the processor. First, the raw
-Wikipedia document dumps need to be transformed into plain text. At the same
-step, the formulas need to be evaluated and replaced by place-holder words. We
-used the \emph{WikiText} framework of the Mylyn project with a custom builder
-to transform the document into plain text. For analysing the \TeX \ mark-up in
-the \texttt{<math/>} tags, \emph{SnuggleTeX} transforms the formulas into
-their MathML representations, where variable names can be easily extracted.
-
-The second computation heavy part is tokenization and tagging of the plain
-text documents. We used the Stanford maximum entropy tagger with the \emph{WSJ
-left3words} language model.
-
-Finally, the rating for all nouns, contained in all candidate sentences, needs
-to be calculated.
+\paragraph{Implementation}
+We implemented the MLP processing system \cite{github} as Stratosphere data-flow using JAVA which allows for scalable execution, application of complex higher order functions and easy integration of third party tool like Stanford NLP and the Mylyn framework.
+%
+%In contrast to MapReduce that supports only the second order functions map and reduce, we used the CoGroup function 
+%
+%\tikzset{actor/.style={
+%        rectangle,
+%        minimum size=6mm,
+%        very thick,
+%        draw=gray!50!black!50,
+%        top color=white,
+%        bottom color=gray!50!black!20
+%    },
+%    arrow/.style={
+%        -latex, thick, shorten <=2pt,shorten >=2pt
+%    }
+%}
+%\begin{figure}[H]
+%	\begin{tikzpicture}[node distance=5mm and 8mm]
+%		\node (Input) [align=center]{Wiki Dump};
+%		\node (DocumentParser) [actor, right=of Input, align=center] {\emph{Map}\\\textbf{Parser}};
+%		\node (Candidates) [actor, right=of DocumentParser, align=center] {\emph{CoGroup}\\\textbf{Kernel}};
+%		\node (Sentence) [actor, above=of DocumentParser, align=center] {\emph{Map}\\\textbf{Tagger}};
+%		\node (Filter) [actor, right=of Candidates, align=center] {\emph{Reduce}\\\textbf{Filter}};
+%		\node (Output) [right=of Filter,align=center]{Raw\ Candidates};
+%		\draw[arrow] (Input)--(DocumentParser);
+%		\draw[arrow] (DocumentParser)--(Sentence);
+%		\draw[arrow] (DocumentParser)--(Candidates);
+%		\draw[arrow] (Sentence.east)--(Candidates.north);
+%		\draw[arrow] (Candidates)--(Filter);
+%		\draw[arrow] (Filter)--(Output);
+%	\end{tikzpicture}
+%\caption{Data flow of the Stratosphere program: We used an
+%ordered list of tokenized sentences and a list of identified variables, to
+%compute the ranking score. A \emph{CoGroup Contract} takes both lists and
+%computes the subsets, build upon distinct document keys. Finally, a
+%\emph{Reduce Contract} filters out results below a given threshold $\theta$.}
+%\end{figure}
+%
+%There are three computation heavy parts in the processor. First, the raw
+%Wikipedia document dumps need to be transformed into plain text. At the same
+%step, the formulas need to be evaluated and replaced by place-holder words. We
+%used the \emph{WikiText} framework of the Mylyn project with a custom builder
+%to transform the document into plain text. For analyzing the \TeX \ mark-up in
+%the \texttt{<math/>} tags, \emph{SnuggleTeX} transforms the formulas into
+%their MathML representations, where variable names can be easily extracted.
+%
+%The second computation heavy part is tokenization and tagging of the plain
+%text documents. We used the Stanford maximum entropy tagger with the \emph{WSJ
+%left3words} language model.
+%
+%Finally, the rating for all nouns, contained in all candidate sentences, needs
+%to be calculated.
+%
 
@@
 \section{Evaluation}
-%\subsubsection{Variable Retrieval}
-\label{vr}
-Throughout our experiments we made some observations that had an impact on the
-accuracy of retrieving the correct set of variables. First of all, people tend
-to incorrectly use \TeX\ trying to create formulas. E.g.,
-\texttt{\textbackslash text\{log\}} is more often used than the correct
-operator \texttt{\textbackslash log}. Another problem is that sometimes people
-use indices as a form of annotation, like $T_{before}$, $R_{A}$, and $R_{B}$.
-These ambiguities lead to a increased number of false positives.
-
-We took a very conservative approach and preprocessed all formulas.
-\texttt{\textbackslash text\{\}} blocks along with plain subscriptions will be
-removed before analysis. Moreover, we created a comprehensive blacklist to
-improve the results further. Short words like `a' and `i', which are very
-common, could be easily matched by our processor if a formula contains one of
-those symbols. Additionally, we blacklist common mathematical operators,
-constants, and functions.
+We extracted 460 identifiers definition tuples in 550000 relations with a
+high score ($>0.8$). The most common identifier definition tuples were 
+number $n$ (1709), time $t$	(1480), mass $M$ (1042), radius $r$ (752), temperature	$T$	(666), angle $\theta$ (639), group $G$ (635).
+%\todo{I don't trust this evaluation and I'm trying to rewrite that}
+%%\subsubsection{Variable Retrieval}
+%\label{vr}
+%Throughout our experiments we made some observations that had an impact on the
+%accuracy of retrieving the correct set of variables. First of all, people tend
+%to incorrectly use \TeX\ trying to create formulas. E.g.,
+%\texttt{\textbackslash text\{log\}} is more often used than the correct
+%operator \texttt{\textbackslash log}. Another problem is that sometimes people
+%use indices as a form of annotation, like $T_{before}$, $R_{A}$, and $R_{B}$.
+%These ambiguities lead to a increased number of false positives.
+%
+%We took a very conservative approach and preprocessed all formulas.
+%\texttt{\textbackslash text\{\}} blocks along with plain subscriptions will be
+%removed before analysis. Moreover, we created a comprehensive blacklist to
+%improve the results further. Short words like `a' and `i', which are very
+%common, could be easily matched by our processor if a formula contains one of
+%those symbols. Additionally, we blacklist common mathematical operators,
+%constants, and functions.
 
@@
 
+\section{Further work}
+Our first intuition was to discover grammatical patters like \emph{let \texttt{variable} be \texttt{definition}} or were \emph{\texttt{variable} (indicates/stands for/ denotes)\texttt{definition}} based on the statistical findings.
+However, our impression is now that the performance gain would bi little.
+
+Now, we are thinking to improve the distance measure between identifier and definition by considering the grammatical structure of the sentence.
+I.e the amount of distance added to scoring function should depend on the grammatical type of the relation between the words.
+
+To improve the robustness of the \emph{term frequency} measure further, one
+might cluster sets of scientific documents, based on their specific field of
+research.
+
 \section{Conclusion}
@@
 about the text surface, can lead to quality results. However, this approach is
-only applicable under certain conditions. In situations where both parts of a
-relation are unknown, other methods, especially supervised ones, are used to
-bootstrap a language model. Unfortunately, we still need a labelled test
+only applicable under certain conditions. 
+If there is only few data for some identifier our statistical approach tends to fail. In that situations other methods, especially supervised ones, are used to
+bootstrap a language model. Unfortunately, we still need a labeled test
 corpus to measure the performance of a classifier, trained with our generated
@@
 
-Apart from the mentioned weighting parameters, the ranking's quality is
-influenced by two further parameters. These are the assumptions about the
-parameter $c$. Balancing all these parameters requires extensive knowledge
-about the underlying language. These parameters might be specific for every
-type of relation. Therefore, this strategy depends heavily on human input and
-many iterations of tuning.
+Furthermore it would be nice to have a systematic approach for a wise choice of the ranking parameters.
 
-To further improve the robustness of the \emph{term frequency} measure, one
-might cluster sets of scientific documents, based on their specific field of
-research.
-
-%460 identifiers
 \begingroup

### 2014-03-13 paper.tex
Further improvements
Thanks to @HowardCohl for the suggestions

@@
 
-\institute{Berlin Institute of Technology, Database Systems and Information Management Group,\\
+\institute{TU Berlin, Database Systems and Information Management Group,\\
 Einsteinufer 17, 10587 Berlin, Germany\\
 \mailsa\\
-\url{http://www.dima.tu-berlin.de/}}
+\url{http://demo.formulasearchengine.com/}}
 
@@
 \begin{abstract}
-In natural language words and phrases themselves imply the semantics.
-Complimentary, the meaning of identifiers in mathematical formulae is undefined.
-Thus scientist must study the context to decode the meaning.
+In natural language, words and phrases themselves imply the semantics.
+In contrast, the meaning of identifiers in mathematical formulae is undefined.
+Thus scientists must study the context to decode the meaning.
 The \emph{Mathematical Language Processing} project, aims to support that process and
 presents automated identifier definition discovery.
-
 Ordinary natural language processing approaches are trained on large, manually annotated news
-corpora. They fall short in reasonable quality of definition discovery for scientific texts.
+corpora.
+They fall short in reasonable quality of definition discovery for scientific texts.
 We use statistical methods to discover definitions in the surrounding text.
-
 The evaluation of our prototypical system, applied on the Wikipedia text
 corpus shows that our approach augments the user experience
-substantially. While hovering the identifiers occurring
-in the formula tool-tips with the most probable definitions
+substantially.
+While hovering the identifiers in the formula, tool-tips with the most probable definitions
 occur.
-Tests with random samples show, that the displayed definitions provide
+Tests with random samples show that the displayed definitions provide
 a good match with the actual meaning of the identifiers.   
@@
 Mathematical formulae are a viable source of information for a wide range of
-scientists. Often, they contain variables that are unknown or ambiguous to the reader.
+scientists.
+Often, they contain variables that are unknown or ambiguous to the reader.
 Therefore, one needs to study the surrounding text to find the relevant definition.
+An automatic information retrieval system can be used to reduce the reader's
+effort.
+Students and scientists of other disciplines would especially profit from a 
+system that helps them to understand formulae.
 
-An automatic information retrieval system can be used to reduce the readers
-effort. Especially students and scientists of other
-disciplines would profit from a system that helps them to understand formulae.
-
-To build such a system a labeled text corpus that annotates variables
-and their definition is desirable. At the project start, such a corpus was not available.
-%Unfortunately, no one has built . Instead of labeling large amounts of text by hand,
+To build such a system, a labeled text corpus that annotates variables
+and their definition is desirable.
+At the project start, such a corpus was not available.
 Consequently, we use simple statistical assumptions about co-occurrences between identifiers
-and their potential definition. In combination with the Stratosphere system
-that is capable of scalable data processing,  we are able to generate a large amount of 
-definition relation candidates in a reasonable and a little implementation effort.
-% the data processing system Stratosphere we are able to analyze large text corpora   and a system that  investigate a potential way of automatically labeling large corpora with a set
-%of statistical assumptions.
+and their potential definition.
+In combination with the Stratosphere system
+that is capable of scalable data processing,  we generate a large amount of 
+definition relation 
+candidates fast with only minimal implementation overhead to ensure the scalability of our algorithm.
 
@@
 
-The English Wikipedia contains roughly four million articles. Even if we only
-pick articles containing \texttt{<math/>} tag, our processor still needs
-to compute tens of thousands of articles. Especially when using a maximum
+The English Wikipedia contains roughly four million articles.
+Even if we only
+pick articles containing the \texttt{<math/>} tag, our processor still needs
+to compute with tens of thousands of articles.
+Especially when using a maximum
 entropy \emph{POS} tagger \cite{Rathna96}, like the one in Stanford's NLP
@@
 \paragraph{Related Work}
-\citeauthor{Quoc2010}\cite{Quoc2010} proposed an approach for
+\citeauthor{Quoc2010} \cite{Quoc2010} proposed an approach for
 relating whole formulas to sentences and their describing paragraphs.
-\citeauthor{Yokoi}\cite{Yokoi} trained a \emph{support vector machine} to extract
+\citeauthor{Yokoi} \cite{Yokoi} trained a \emph{support vector machine} to extract
 natural language descriptions for mathematical expressions.
@@
 we assume that the identifier and the definition co-occur in the same sentence.
-Furthermore, we restrict ourselves to identifiers that co-occur with a noun
-phrase in one sentence in the document at least.
-
-For each description candidate we use the weighted sum
+Each description candidate is scored with the weighted sum
 \begin{equation} \label{eq:rating}
@@
 +\beta{R}_{\sigma_\mathrm s}(n)
-		+\gamma\mathrm{tf}(t,d)}{\alpha+\beta+\gamma} \mapsto [0;1]
++\gamma\mathrm{tf}(t,d)}{\alpha+\beta+\gamma} \mapsto [0;1].
 \end{equation}
-that respects the distance $\Delta$ between identifier and definition term $t$, the distance $n$ between the display-style equation and th sentence that contains the term and the identifier and the term frequency $\mathrm{tf}(t,d)$ in the current document $d$
-with a normal distributed distance measure $R_\sigma(\Delta)= \exp\left[-\frac{1}{2}\frac{\Delta^2-1}{\sigma^2}\right].$
-Our model assumption is that the probability to find a relation at $\Delta=1$ like for example in \emph{the energy $E$ and the mass $m$} is maximal, i.e.
-$R_{\sigma_\mathrm d}(1)=1$.
-In order to determine the full width half maximum of our distribution we evaluated some articles manually and found $R_{\sigma_\mathrm d}(1)\approx 2 R_{\sigma_\mathrm d}(5)$ and thus $\sigma_d=\sqrt\frac{5^2-1}{2\ln 2}$.
+It respects the distance $\Delta$ between identifier and definition term $t$, the distance $n$ between the display-style equation and the sentence that contains the term and the identifier and the term frequency $\mathrm{tf}(t,d)$ in the current document $d$.
+The distance was normalized with $R_\sigma(\Delta)= \exp\left[-\frac{1}{2}\frac{\Delta^2-1}{\sigma^2}\right].$
+We assume that the probability to find a relation at $\Delta=1,$ is maximal.
+For example in the text fragment \emph{the energy $E$ and the mass $m$} $\Delta(E,\textsf{energy})=\Delta(m,\textsf{mass})=1$ and $R_{\sigma_\mathrm d}(1)=1.$
+In order to determine the full width at half maximum of our distribution we evaluated some articles manually and found $R_{\sigma_\mathrm d}(1)\approx 2 R_{\sigma_\mathrm d}(5)$ and thus 
+$\sigma_d=\sqrt\frac{12}{\ln 2}$.
 
-The probability to find a correct definition decays for 50 \% within three sentences. Consequently  $\sigma_\mathrm s=2\left({\ln 2}\right)^{-\frac{1}{2}}$.
+The probability to find a correct definition decays for 50\% within three sentences.
+Consequently  $\sigma_\mathrm s=2\left({\ln 2}\right)^{-\frac{1}{2}}$.
 
 The classic tf-idf \cite{Salton86} statistic reflects the importance of a term to a document.
-For our task the inverse document frequency (idf)  would assign high penalties to frequent which are valid definitions for identifiers like in the example `Energy` and `Einstein`.
-Consequently we used just the term frequency which still had an impact on the overall result that seemed to high in our manual experiments.
+For our task the inverse document frequency (idf)  would assign high penalties to frequent which are valid definitions for identifiers like in the example `Energy' and `Einstein'.
+Consequently we used just the term frequency which still had an impact on the overall result 
+that seemed too high in our manual experiments.
 Therefore, we use the tuning parameters $\gamma=0.75$ and $\alpha = \beta = 1$.
-
-%The weighting parameters $\alpha$, $\beta$, and $\gamma$ in Equation
-%\ref{eq:rating} are used to balance the contribution of all embedded
-%statistics to the overall ranking. But these parameters need to be tuned
-%carefully. Otherwise the results tend to be biased towards one parameter. In
-%other words, if the distance measure $R_{D}$ dominates the other measures,
-%more complex\footnote{in terms of words between both entities, participating
-%in a relation} relations are less likely to be discovered.
-%
-%The \emph{term frequency} turns out to be the least robust statistic. Usually,
-%the frequency of a specific term co-occurring with a variable name within one
-%document is rather low. Thus, the derived ranking is relatively volatile. In
-%our experiments, we chose $\gamma$ to be 0.75 ($\alpha = \beta = 1$) for a
-%more robust ranking.
-%to  E.g., `Hamiltonian' as well as `length' are both valid entities. Obviously, the number of documents containing one of these words, within a corpus like Wikipedia, differs a lot.
-%
-%Therefore, we only use the normalized term frequency. $t$ and $w$ are terms.
-%$f(t,d)$ denotes the frequency of a term within a document $d$.
-%
-%\begin{equation}
-%	\label{eq:tf}
-%	\mathrm{tf}(t,d) = \frac{\mathrm{f}(t,d)}{\max\{\mathrm{f}(w,d):w \in d\}}
-%\end{equation}
-%
 
@@
 We implemented the MLP processing system \cite{github} as Stratosphere data-flow using JAVA which allows for scalable execution, application of complex higher order functions and easy integration of third party tool like Stanford NLP and the Mylyn framework.
-%
-%In contrast to MapReduce that supports only the second order functions map and reduce, we used the CoGroup function 
-%
-%\tikzset{actor/.style={
-%        rectangle,
-%        minimum size=6mm,
-%        very thick,
-%        draw=gray!50!black!50,
-%        top color=white,
-%        bottom color=gray!50!black!20
-%    },
-%    arrow/.style={
-%        -latex, thick, shorten <=2pt,shorten >=2pt
-%    }
-%}
-%\begin{figure}[H]
-%	\begin{tikzpicture}[node distance=5mm and 8mm]
-%		\node (Input) [align=center]{Wiki Dump};
-%		\node (DocumentParser) [actor, right=of Input, align=center] {\emph{Map}\\\textbf{Parser}};
-%		\node (Candidates) [actor, right=of DocumentParser, align=center] {\emph{CoGroup}\\\textbf{Kernel}};
-%		\node (Sentence) [actor, above=of DocumentParser, align=center] {\emph{Map}\\\textbf{Tagger}};
-%		\node (Filter) [actor, right=of Candidates, align=center] {\emph{Reduce}\\\textbf{Filter}};
-%		\node (Output) [right=of Filter,align=center]{Raw\ Candidates};
-%		\draw[arrow] (Input)--(DocumentParser);
-%		\draw[arrow] (DocumentParser)--(Sentence);
-%		\draw[arrow] (DocumentParser)--(Candidates);
-%		\draw[arrow] (Sentence.east)--(Candidates.north);
-%		\draw[arrow] (Candidates)--(Filter);
-%		\draw[arrow] (Filter)--(Output);
-%	\end{tikzpicture}
-%\caption{Data flow of the Stratosphere program: We used an
-%ordered list of tokenized sentences and a list of identified variables, to
-%compute the ranking score. A \emph{CoGroup Contract} takes both lists and
-%computes the subsets, build upon distinct document keys. Finally, a
-%\emph{Reduce Contract} filters out results below a given threshold $\theta$.}
-%\end{figure}
-%
-%There are three computation heavy parts in the processor. First, the raw
-%Wikipedia document dumps need to be transformed into plain text. At the same
-%step, the formulas need to be evaluated and replaced by place-holder words. We
-%used the \emph{WikiText} framework of the Mylyn project with a custom builder
-%to transform the document into plain text. For analyzing the \TeX \ mark-up in
-%the \texttt{<math/>} tags, \emph{SnuggleTeX} transforms the formulas into
-%their MathML representations, where variable names can be easily extracted.
-%
-%The second computation heavy part is tokenization and tagging of the plain
-%text documents. We used the Stanford maximum entropy tagger with the \emph{WSJ
-%left3words} language model.
-%
-%Finally, the rating for all nouns, contained in all candidate sentences, needs
-%to be calculated.
-%
 
-%\subsection{Experiments}
 \section{Evaluation}
 We extracted 460 identifiers definition tuples in 550000 relations with a
-high score ($>0.8$). The most common identifier definition tuples were 
+high score ($>0.8$).
+The most common identifier definition tuples were 
 number $n$ (1709), time $t$	(1480), mass $M$ (1042), radius $r$ (752), temperature	$T$	(666), angle $\theta$ (639), group $G$ (635).
-%\todo{I don't trust this evaluation and I'm trying to rewrite that}
-%%\subsubsection{Variable Retrieval}
-%\label{vr}
-%Throughout our experiments we made some observations that had an impact on the
-%accuracy of retrieving the correct set of variables. First of all, people tend
-%to incorrectly use \TeX\ trying to create formulas. E.g.,
-%\texttt{\textbackslash text\{log\}} is more often used than the correct
-%operator \texttt{\textbackslash log}. Another problem is that sometimes people
-%use indices as a form of annotation, like $T_{before}$, $R_{A}$, and $R_{B}$.
-%These ambiguities lead to a increased number of false positives.
-%
-%We took a very conservative approach and preprocessed all formulas.
-%\texttt{\textbackslash text\{\}} blocks along with plain subscriptions will be
-%removed before analysis. Moreover, we created a comprehensive blacklist to
-%improve the results further. Short words like `a' and `i', which are very
-%common, could be easily matched by our processor if a formula contains one of
-%those symbols. Additionally, we blacklist common mathematical operators,
-%constants, and functions.
-
-Because of the fact, that we do not have any annotated test corpora,
-evaluation has to be performed by hand. Therefore, we took a sample of 30
-random documents and counted all matches. The resulting estimates for
+Because we do not have any annotated test corpora,
+evaluation has to be performed by hand.
+Therefore, we took a sample of 30
+random documents and counted all matches.
+The resulting estimates for
 \emph{recall} and \emph{precision} are 0.99 and 0.86 respectively.
@@
 \section{Further work}
-Our first intuition was to discover grammatical patters like \emph{let \texttt{variable} be \texttt{definition}} or were \emph{\texttt{variable} (indicates/stands for/ denotes)\texttt{definition}} based on the statistical findings.
-However, our impression is now that the performance gain would bi little.
+Our first intuition was to discover grammatical patters 
+like \emph{let \texttt{variable} be \texttt{definition}} or 
+were \emph{\texttt{variable} (indicates/stands for/ denotes)\texttt{definition}} based on the statistical findings.
+However, our impression is now that the performance gain would \todo{``be little'' doesn't read good.}be little.
 
-Now, we are thinking to improve the distance measure between identifier and definition by considering the grammatical structure of the sentence.
-I.e the amount of distance added to scoring function should depend on the grammatical type of the relation between the words.
+Now we are thinking to improve the distance measure between identifier and definition by 
+considering the grammatical structure of the sentence,
+i.e., the amount of distance added to the scoring function should depend on the grammatical 
+type of the relation between the words.
 
@@
 Our experiments showed that combining a POS tagger with numerical statistics
-about the text surface, can lead to quality results. However, this approach is
+about the text surface, can lead to quality results.
+However, this approach is
 only applicable under certain conditions. 
-If there is only few data for some identifier our statistical approach tends to fail. In that situations other methods, especially supervised ones, are used to
-bootstrap a language model. Unfortunately, we still need a labeled test
+For seldom identifiers our statistical approach tends to fail.
+In that situation, other methods, especially supervised ones, are used to
+bootstrap a language model.
+Unfortunately, we still need a labeled test
 corpus to measure the performance of a classifier, trained with our generated
@@
 
-Furthermore it would be nice to have a systematic approach for a wise choice of the ranking parameters.
+Furthermore, a systematic approach for determining a wise choice of the ranking parameters might improve the overall performance of our system a lot.
+\paragraph*{Acknowledgment}
+Thanks to Howard Cohl for proof reading the paper and to the course instructor of the database project course in fall 2012 Holmer Hemsen.
+The implementation and a first draft of the this papers was completed in the scope of this course.
 

### 2014-03-13 paper.tex
screenshot, references and corrections
thanks again to @HowardCohl for proof reading

@@
 \usepackage{todonotes}
+\usepackage{wrapfig}
 
@@
 The evaluation of our prototypical system, applied on the Wikipedia text
-corpus shows that our approach augments the user experience
+corpus, shows that our approach augments the user experience
 substantially.
@@
 Often, they contain variables that are unknown or ambiguous to the reader.
-Therefore, one needs to study the surrounding text to find the relevant definition.
+Therefore, one usually needs to study the surrounding text to find the relevant definition.
 An automatic information retrieval system can be used to reduce the reader's
@@
 In combination with the Stratosphere system
-that is capable of scalable data processing,  we generate a large amount of 
+that is capable of scalable data processing,  we rapidly generate a large amount of 
 definition relation 
-candidates fast with only minimal implementation overhead to ensure the scalability of our algorithm.
+candidates with only minimal implementation overhead to ensure the scalability of our algorithm.
 
+%\pagebreak
+\begin{wrapfigure}{r}{0.5\textwidth}
+	\label{fig:screenshot}
+  \includegraphics[width=0.5\textwidth]{screenshot}
+  \caption{Screen-shot of the energy mass relation page `Mass–energy equivalence`, while hovering the letter `E`.}
+  \vspace{-20pt}
+\end{wrapfigure}
 We chose the Wikipedia text corpus as a target because of two facts.
-First, it is easy to extract the formulas and identifiers from the text,
-since the input language for formula texvc is very restrictive.
-Second, we can leverage the semantic information between mark-up and
+First, because the input language for formula texvc is very restrictive, it is easy to extract the formulas and identifiers from the text.
+Second, we are able to leverage the semantic information between mark-up and
 word tokens, e.g., hyperlinks.
@@
 computation.
-
 Therefore we implement the proposed strategy with the Stratosphere system based
@@
 \paragraph{Related Work}
+
 \citeauthor{Quoc2010} \cite{Quoc2010} proposed an approach for
@@
 \section{Statistical definition discovery}
+
 We detect relations between identifiers and their description in two steps.
-First, we extract the identifiers from the display-style formulae and
+First, we extract the identifiers from the formulae and
 second we determine their description from the surrounding text.
@@
 \end{equation}
-It respects the distance $\Delta$ between identifier and definition term $t$, the distance $n$ between the display-style equation and the sentence that contains the term and the identifier and the term frequency $\mathrm{tf}(t,d)$ in the current document $d$.
+It respects the distance $\Delta$ between identifier and definition term $t$, the distance $n$ between the display-style equation and the sentence that contains the term, the identifier and the term frequency $\mathrm{tf}(t,d)$ in the current document $d$.
 The distance was normalized with $R_\sigma(\Delta)= \exp\left[-\frac{1}{2}\frac{\Delta^2-1}{\sigma^2}\right].$
 We assume that the probability to find a relation at $\Delta=1,$ is maximal.
-For example in the text fragment \emph{the energy $E$ and the mass $m$} $\Delta(E,\textsf{energy})=\Delta(m,\textsf{mass})=1$ and $R_{\sigma_\mathrm d}(1)=1.$
+For example in the text fragment \emph{the energy $E$, and the mass $m$} $\Delta(E,\textsf{energy})=\Delta(m,\textsf{mass})=1$ and $R_{\sigma_\mathrm d}(1)=1.$
 In order to determine the full width at half maximum of our distribution we evaluated some articles manually and found $R_{\sigma_\mathrm d}(1)\approx 2 R_{\sigma_\mathrm d}(5)$ and thus 
 $\sigma_d=\sqrt\frac{12}{\ln 2}$.
-
 The probability to find a correct definition decays for 50\% within three sentences.
@@
 The classic tf-idf \cite{Salton86} statistic reflects the importance of a term to a document.
-For our task the inverse document frequency (idf)  would assign high penalties to frequent which are valid definitions for identifiers like in the example `Energy' and `Einstein'.
-Consequently we used just the term frequency which still had an impact on the overall result 
-that seemed too high in our manual experiments.
-Therefore, we use the tuning parameters $\gamma=0.75$ and $\alpha = \beta = 1$.
+For our task, the inverse document frequency (idf) would assign high penalties to frequent words like `energy` in comparison to seldom words like `Einstein` which are valid definitions for identifiers.
+As the influence of $\mathrm{tf}(t,d)$ on the overall ranking \ref{eq:rating}, seems to be very high, we reduce the impact with the tuning parameters $\gamma=0.75$ and $\alpha = \beta = 1$.
 
 \paragraph{Implementation}
-We implemented the MLP processing system \cite{github} as Stratosphere data-flow using JAVA which allows for scalable execution, application of complex higher order functions and easy integration of third party tool like Stanford NLP and the Mylyn framework.
-
+We implemented the MLP processing system \cite{github} as Stratosphere data-flow using JAVA which allows for scalable execution, application of complex higher order functions and easy integration of third party tools like Stanford NLP and the Mylyn framework.
 \section{Evaluation}
@@
 high score ($>0.8$).
+
 The most common identifier definition tuples were 
@@
 random documents and counted all matches.
-The resulting estimates for
-\emph{recall} and \emph{precision} are 0.99 and 0.86 respectively.
+Most of the retrieved definitions were considered as relevant.
+Furthermore, for more than 50 \% of the investigated identifiers our system detects are relevant definition.
+However, we neither consider identifiers with subscripts nor identifiers that do not occur in the surrounding text.  
+%The resulting estimates for
+%\emph{recall} and \emph{precision} are 0.99 and 0.86 respectively.
 
@@
 were \emph{\texttt{variable} (indicates/stands for/ denotes)\texttt{definition}} based on the statistical findings.
-However, our impression is now that the performance gain would \todo{``be little'' doesn't read good.}be little.
+However, our impression is that there would not be significant performance gain.
 
-Now we are thinking to improve the distance measure between identifier and definition by 
-considering the grammatical structure of the sentence,
-i.e., the amount of distance added to the scoring function should depend on the grammatical 
-type of the relation between the words.
+The distance measure
+$R_{\sigma_d}$
+fails for the example of figure \ref{fig:screenshot} since
+$\Delta(\mathsf{energy},E)=\Delta(\mathsf{mass},m)=2$.
+We plan to consider the grammatical structure of the sentence for an improved distance measure $\Delta^{+} : \Delta^+(\mathsf{energy},E)\ge\Delta^+(\mathsf{mass},m)$.
+Currently, we use the score $R$ to identify the most probably definition-identifier tuple on each page, even it occurs multiple times on the page.
+%((E|e)nergy.{1,100}''E''|''E''.{1,100}(E|e)nergy)
+For the example the relation `energy-$E$` was found 21 times by our system on the page.
+A promising approach, is to use $R^\Sigma=\sum_{i=1}^n 2^{-i} R_i $ where $R_i$ is a sorted list i.e. $R_1$ the highest ranked definition for that relation according to the current measure $R$.
 
 To improve the robustness of the \emph{term frequency} measure further, one
-might cluster sets of scientific documents, based on their specific field of
+might cluster sets of scientific documents, based on their specific fields of
 research.
@@
 bootstrap a language model.
-Unfortunately, we still need a labeled test
-corpus to measure the performance of a classifier, trained with our generated
-data.
-
-Furthermore, a systematic approach for determining a wise choice of the ranking parameters might improve the overall performance of our system a lot.
+Unfortunately, they require a labeled test
+corpus to measure the performance of a classifier that could be trained with our generated data.
+Currently, we are investigating a cooperation with the group of professor \citeauthor{Aizawa2013} and the use of the NTCIR-Math-10 Task, Math Understanding Subtask gold standard dataset \cite{Aizawa2013} for further experiments.
+Furthermore, a systematic approach for determining a wise choice of the ranking parameters should significantly improve the overall performance of our system.
 \paragraph*{Acknowledgment}
-Thanks to Howard Cohl for proof reading the paper and to the course instructor of the database project course in fall 2012 Holmer Hemsen.
-The implementation and a first draft of the this papers was completed in the scope of this course.
+Thanks to Howard Cohl for proof reading the paper and to the course instructor of the database project course in fall 2012, Holmer Hemsen.
+The implementation and a first draft of this paper was completed the duration of this course.
 

### 2014-03-13 paper.tex
Further improvements
Thanks to @HowardCohl

@@
 \usepackage{float}
-\usepackage{todonotes}
 \usepackage{wrapfig}
@@
 
-\institute{TU Berlin, Database Systems and Information Management Group,\\
+\institute{
+Database Systems and Information Management Group,\\
+Technische Universit\"{a}t Berlin,
 Einsteinufer 17, 10587 Berlin, Germany\\
@@
 Thus scientists must study the context to decode the meaning.
-The \emph{Mathematical Language Processing} project, aims to support that process and
+The \emph{Mathematical Language Processing} project aims to support that process and
 presents automated identifier definition discovery.
@@
 At the project start, such a corpus was not available.
-Consequently, we use simple statistical assumptions about co-occurrences between identifiers
-and their potential definition.
+We use simple statistical assumptions about co-occurrences between identifiers
+and their potential definition to extract the meaning of the identifiers from their context.
 In combination with the Stratosphere system
@@
   \includegraphics[width=0.5\textwidth]{screenshot}
-  \caption{Screen-shot of the energy mass relation page `Mass–energy equivalence`, while hovering the letter `E`.}
+  \caption{Screenshot of the energy mass relation page `Mass–energy equivalence`, while hovering the letter `E`.}
   \vspace{-20pt}
@@
 
-\paragraph{Related Work}
+\paragraph{Related Work.}
 
@@
 \end{equation}
-It respects the distance $\Delta$ between identifier and definition term $t$, the distance $n$ between the display-style equation and the sentence that contains the term, the identifier and the term frequency $\mathrm{tf}(t,d)$ in the current document $d$.
+It depends on the distance $\Delta$ between identifier and definition term $t$, the distance $n$ between the display-style equation and the sentence that contains the term, and the identifier and the term frequency $\mathrm{tf}(t,d)$ in the current document $d$.
 The distance was normalized with $R_\sigma(\Delta)= \exp\left[-\frac{1}{2}\frac{\Delta^2-1}{\sigma^2}\right].$
-We assume that the probability to find a relation at $\Delta=1,$ is maximal.
-For example in the text fragment \emph{the energy $E$, and the mass $m$} $\Delta(E,\textsf{energy})=\Delta(m,\textsf{mass})=1$ and $R_{\sigma_\mathrm d}(1)=1.$
-In order to determine the full width at half maximum of our distribution we evaluated some articles manually and found $R_{\sigma_\mathrm d}(1)\approx 2 R_{\sigma_\mathrm d}(5)$ and thus 
+We assume that the probability to find a relation at $\Delta=1$ is maximal.  For example in the 
+text fragment \emph{the energy $E$, and the mass $m$,}  
+In order to determine the full width at half maximum of our distribution, we evaluated some articles manually and found $R_{\sigma_\mathrm d}(1)\approx 2 R_{\sigma_\mathrm d}(5)$ and thus 
 $\sigma_d=\sqrt\frac{12}{\ln 2}$.
@@
 The classic tf-idf \cite{Salton86} statistic reflects the importance of a term to a document.
-For our task, the inverse document frequency (idf) would assign high penalties to frequent words like `energy` in comparison to seldom words like `Einstein` which are valid definitions for identifiers.
-As the influence of $\mathrm{tf}(t,d)$ on the overall ranking \ref{eq:rating}, seems to be very high, we reduce the impact with the tuning parameters $\gamma=0.75$ and $\alpha = \beta = 1$.
+For our task, the inverse document frequency (idf) would assign high penalties to 
+frequent words like `energy', in comparison to 
+%\todo{I replaced `seldom words' with `words seldom seen'.  Is that ok?}
+words seldom seen like `Einstein', which are valid definitions for identifiers.
+As the influence of $\mathrm{tf}(t,d)$ on the overall ranking (\ref{eq:rating}) seems to be very high, we reduce the impact with the tuning parameters $\gamma=0.75$ and $\alpha = \beta = 1$.
 
-\paragraph{Implementation}
-We implemented the MLP processing system \cite{github} as Stratosphere data-flow using JAVA which allows for scalable execution, application of complex higher order functions and easy integration of third party tools like Stanford NLP and the Mylyn framework.
+\paragraph{Implementation.}
+We implemented the MLP processing system \cite{github} as Stratosphere data-flow using Java which allows for scalable execution, application of complex higher order functions and easy integration of third party tools like Stanford NLP and the Mylyn framework.
 \section{Evaluation}
 We extracted 460 identifiers definition tuples in 550000 relations with a
-high score ($>0.8$).
-
-The most common identifier definition tuples were 
-number $n$ (1709), time $t$ (1480), mass $M$ (1042), radius $r$ (752), temperature	$T$	(666), angle $\theta$ (639), group $G$ (635).
+high score ($>0.8$).  The most common identifier definition tuples were 
+number $n$ (1709), time $t$ (1480), mass $M$ (1042), radius $r$ (752), 
+temperature $T$ (666), angle $\theta$ (639), group $G$ (635).
 Because we do not have any annotated test corpora,
@@
 Most of the retrieved definitions were considered as relevant.
-Furthermore, for more than 50 \% of the investigated identifiers our system detects are relevant definition.
-However, we neither consider identifiers with subscripts nor identifiers that do not occur in the surrounding text.  
+Furthermore, for more than 50\% of the investigated identifiers, 
+the most probable meanings detected by our system were relevant.
+However, we neither consider identifiers with subscripts, nor identifiers that do not occur in the surrounding text.  
 %The resulting estimates for
@@
 \section{Further work}
-Our first intuition was to discover grammatical patters 
+Our original intuition was to discover grammatical patters 
 like \emph{let \texttt{variable} be \texttt{definition}} or 
 were \emph{\texttt{variable} (indicates/stands for/ denotes)\texttt{definition}} based on the statistical findings.
-However, our impression is that there would not be significant performance gain.
+However, our impression is that this 
+would not lead to significant performance gain.
 
@@
 $R_{\sigma_d}$
-fails for the example of figure \ref{fig:screenshot} since
+fails for the example of Fig.~\ref{fig:screenshot} since
 $\Delta(\mathsf{energy},E)=\Delta(\mathsf{mass},m)=2$.
 We plan to consider the grammatical structure of the sentence for an improved distance measure $\Delta^{+} : \Delta^+(\mathsf{energy},E)\ge\Delta^+(\mathsf{mass},m)$.
-Currently, we use the score $R$ to identify the most probably definition-identifier tuple on each page, even it occurs multiple times on the page.
-%((E|e)nergy.{1,100}''E''|''E''.{1,100}(E|e)nergy)
-For the example the relation `energy-$E$` was found 21 times by our system on the page.
-A promising approach, is to use $R^\Sigma=\sum_{i=1}^n 2^{-i} R_i $ where $R_i$ is a sorted list i.e. $R_1$ the highest ranked definition for that relation according to the current measure $R$.
-
+Currently, we use the score $R$ to identify the most probably definition-identifier tuple on each page, even if it occurs multiple times on the page.
+For example, on the `Mass-energy equivalence` page 21 sentences contain the combination of the identifier $E$ and the noun `energy`.
+A promising approach, is to use $R^\Sigma=\sum_{i=1}^n 2^{-i} R_i,$ where $R_i$ is a 
+sorted list, i.e., $R_1$ the highest ranked definition for that relation according to the current measure $R$.
 To improve the robustness of the \emph{term frequency} measure further, one
@@
 only applicable under certain conditions. 
-For seldom identifiers our statistical approach tends to fail.
+For identifiers which
+are seldom seen, our statistical approach tends to fail.
 In that situation, other methods, especially supervised ones, are used to
@@
 corpus to measure the performance of a classifier that could be trained with our generated data.
-Currently, we are investigating a cooperation with the group of professor \citeauthor{Aizawa2013} and the use of the NTCIR-Math-10 Task, Math Understanding Subtask gold standard dataset \cite{Aizawa2013} for further experiments.
+Currently, we are planning a cooperation with the group of professor \citeauthor{Aizawa2013} from the National Institute of Informatics in Japan for further research on that topic.
+In that context we consider to use the NTCIR-Math-10 Task, Math Understanding Subtask gold standard dataset \cite{Aizawa2013} for a comparable evaluation.
 Furthermore, a systematic approach for determining a wise choice of the ranking parameters should significantly improve the overall performance of our system.
-\paragraph*{Acknowledgment}
-Thanks to Howard Cohl for proof reading the paper and to the course instructor of the database project course in fall 2012, Holmer Hemsen.
-The implementation and a first draft of this paper was completed the duration of this course.
+\paragraph*{Acknowledgments.}
+Thanks to Howard Cohl for proof reading the paper and to Holmer Hemsen, the course instructor of the database project course at TU-Berlin in Fall 2012.
+The implementation and a first draft of this paper was completed in the duration of this course.
 

### 2014-03-16 paper.tex
Minor improvements

@@
 \usepackage{float}
+
 \usepackage{wrapfig}
@@
 Ordinary natural language processing approaches are trained on large, manually annotated news
-corpora.
+corpora for instance.
 They fall short in reasonable quality of definition discovery for scientific texts.
@@
 a good match with the actual meaning of the identifiers.   
-\keywords{machine learning, text mining, parallel computing}
+\keywords{definition discovery, text mining, parallel computing}
 \end{abstract}
@@
   \includegraphics[width=0.5\textwidth]{screenshot}
-  \caption{Screenshot of the energy mass relation page `Mass–energy equivalence`, while hovering the letter `E`.}
+  \caption{Screenshot of the energy mass relation page `Mass–energy equivalence', while hovering the letter `E'.}
   \vspace{-20pt}
@@
 text fragment \emph{the energy $E$, and the mass $m$,}  
-In order to determine the full width at half maximum of our distribution, we evaluated some articles manually and found $R_{\sigma_\mathrm d}(1)\approx 2 R_{\sigma_\mathrm d}(5)$ and thus 
+in order to determine the full width at half maximum of our distribution, we evaluated some articles manually and found $R_{\sigma_\mathrm d}(1)\approx 2 R_{\sigma_\mathrm d}(5)$ and thus 
 $\sigma_d=\sqrt\frac{12}{\ln 2}$.
@@
 \section{Further work}
-Our original intuition was to discover grammatical patters 
+Our original intuition was to discover grammatical patterns 
 like \emph{let \texttt{variable} be \texttt{definition}} or 
-were \emph{\texttt{variable} (indicates/stands for/ denotes)\texttt{definition}} based on the statistical findings.
+were \emph{\texttt{variable} (indicates/stands for/denotes) \texttt{definition}} based on the statistical findings.
 However, our impression is that this 
@@
 Currently, we use the score $R$ to identify the most probably definition-identifier tuple on each page, even if it occurs multiple times on the page.
-For example, on the `Mass-energy equivalence` page 21 sentences contain the combination of the identifier $E$ and the noun `energy`.
+For example, on the `Mass-energy equivalence' page, 21 sentences contain the combination of the identifier $E$ and the noun `energy'.
 A promising approach, is to use $R^\Sigma=\sum_{i=1}^n 2^{-i} R_i,$ where $R_i$ is a 
-sorted list, i.e., $R_1$ the highest ranked definition for that relation according to the current measure $R$.
+sorted list, were $R_1$ is the highest ranked definition for that relation according to the current measure $R$.
 To improve the robustness of the \emph{term frequency} measure further, one
@@
 corpus to measure the performance of a classifier that could be trained with our generated data.
-Currently, we are planning a cooperation with the group of professor \citeauthor{Aizawa2013} from the National Institute of Informatics in Japan for further research on that topic.
-In that context we consider to use the NTCIR-Math-10 Task, Math Understanding Subtask gold standard dataset \cite{Aizawa2013} for a comparable evaluation.
+Currently, we are planning to use the
+NTCIR-Math-10 Task, Math Understanding Subtask gold standard dataset \cite{overview}
+for a comparable evaluation.
 Furthermore, a systematic approach for determining a wise choice of the ranking parameters should significantly improve the overall performance of our system.

### 2014-06-01 paper.tex
new version of the paper as well as an identifier/description dataset of two documents

@@
 \begin{abstract}
-In natural language, words and phrases themselves imply the semantics.
-In contrast, the meaning of identifiers in mathematical formulae is undefined.
-Thus scientists must study the context to decode the meaning.
-The \emph{Mathematical Language Processing} project aims to support that process and
-presents automated identifier definition discovery.
-Ordinary natural language processing approaches are trained on large, manually annotated news
-corpora for instance.
-They fall short in reasonable quality of definition discovery for scientific texts.
-We use statistical methods to discover definitions in the surrounding text.
-The evaluation of our prototypical system, applied on the Wikipedia text
-corpus, shows that our approach augments the user experience
-substantially.
-While hovering the identifiers in the formula, tool-tips with the most probable definitions
-occur.
-Tests with random samples show that the displayed definitions provide
-a good match with the actual meaning of the identifiers.   
+
+In natural language, words and phrases themselves imply the semantics. In
+contrast, the meaning of identifiers in mathematical formulae is undefined.
+Thus scientists must study the context to decode the meaning. The
+\emph{Mathematical Language Processing} project aims to support that process
+and presents automated identifier definition discovery. Ordinary natural
+language processing approaches are trained on large, manually annotated text
+corpora. They fall short in reasonable quality of definition discovery for
+scientific texts. We use statistical methods to discover definitions in the
+surrounding text. The evaluation of our prototypical system, applied on the
+Wikipedia text corpus, shows that our approach augments the user experience
+substantially. While hovering the identifiers in the formula, tool-tips with
+the most probable definitions occur. Tests with random samples show that the
+displayed definitions provide a good match with the actual meaning of the
+identifiers.
+
 \keywords{definition discovery, text mining, parallel computing}
@@
 
-Mathematical formulae are a viable source of information for a wide range of
-scientists.
-Often, they contain variables that are unknown or ambiguous to the reader.
-Therefore, one usually needs to study the surrounding text to find the relevant definition.
-An automatic information retrieval system can be used to reduce the reader's
-effort.
-Students and scientists of other disciplines would especially profit from a 
-system that helps them to understand formulae.
-
-To build such a system, a labeled text corpus that annotates variables
-and their definition is desirable.
-At the project start, such a corpus was not available.
-We use simple statistical assumptions about co-occurrences between identifiers
-and their potential definition to extract the meaning of the identifiers from their context.
-In combination with the Stratosphere system
-that is capable of scalable data processing,  we rapidly generate a large amount of 
-definition relation 
-candidates with only minimal implementation overhead to ensure the scalability of our algorithm.
-
-%\pagebreak
-\begin{wrapfigure}{r}{0.5\textwidth}
+\begin{wrapfigure}{r}{0.4\textwidth}
   \label{fig:screenshot}
-  \includegraphics[width=0.5\textwidth]{screenshot}
+  \vspace{-20pt}
+  \includegraphics[width=0.4\textwidth]{screenshot}
   \caption{Screenshot of the energy mass relation page `Mass–energy equivalence', while hovering the letter `E'.}
@@
 \end{wrapfigure}
-We chose the Wikipedia text corpus as a target because of two facts.
-First, because the input language for formula texvc is very restrictive, it is easy to extract the formulas and identifiers from the text.
-Second, we are able to leverage the semantic information between mark-up and
-word tokens, e.g., hyperlinks.
 
-The English Wikipedia contains roughly four million articles.
-Even if we only
-pick articles containing the \texttt{<math/>} tag, our processor still needs
-to compute with tens of thousands of articles.
-Especially when using a maximum
-entropy \emph{POS} tagger \cite{Rathna96}, like the one in Stanford's NLP
+Mathematical formulae are a viable source of information for a wide range of
+scientists. Often, they contain identifier whose meaning might be at first
+unknown or at least ambiguous to the reader (depending on their knownledge).
+Therefore, one usually needs to study the surrounding text to find the
+relevant definition. An automatic information retrieval system can be used to
+reduce the reader's effort by showing the most relevant definition relation
+found to the reader. Students and scientists of other disciplines would
+especially profit from a system that helps them to understand formulae more
+quickly.
+
+To build such a system, a labeled text corpus that annotates identifiers and
+their definition is desirable. At the project start, such a corpus was not
+available. We use simple statistical assumptions about co-occurrences between
+identifiers and their potential description to extract the meaning of the
+identifiers from their context. In combination with the Stratosphere system
+that is capable of scalable data processing,  we rapidly generate a large
+amount of definition relation candidates with only minimal implementation
+overhead to ensure the scalability of our algorithm.
+
+We chose the Wikipedia text corpus as a target because of two facts. First,
+most articles make use of \texttt{<math/>} tags (\emph{texvc} as an input
+language) for formulae. It is easy to detect and the extraction of identifiers
+is well tried. Second, the text is already annotated with mark-up.
+Particularly, hyperlinks to other articles within the Wikipedia are of
+interest as they typically wrap around any number of words and indicate that
+these in combination are relevant in the given context (respectively
+sentence).
+
+The English Wikipedia contains roughly four million articles. Even if we only
+pick articles containing \texttt{<math/>} tags, our processor still needs to
+compute with tens of thousands of articles. Especially when using text
+annotators (e.g. \emph{POS} tagger \cite{Rathna96}), like Stanford's NLP
 framework, one can make use of a parallel processing system to speed up
-computation.
-Therefore we implement the proposed strategy with the Stratosphere system based
-on the PACT programming model \cite{Alexandrov2010}.
+computation. Therefore we implement the proposed strategy with the
+Stratosphere system based on the PACT programming model \cite{Alexandrov2010}.
 
@@
 
-\section{Statistical definition discovery}
+\section{Statistical Definition Discovery}
 
 We detect relations between identifiers and their description in two steps.
-First, we extract the identifiers from the formulae and
+First, we extract the identifiers from the formulae found in an article and
 second we determine their description from the surrounding text.
 
-Inspired by the \emph{Distant Supervision} approach \cite{Mintz2008},
-we assume that the identifier and the definition co-occur in the same sentence.
-Each description candidate is scored with the weighted sum
+Extracting relevant identifiers from the article rely on the assumption that
+the author will use \texttt{<math/>} tags for all formulae. Thus said, a
+formula that is written within a text cannot be recognized, and therefor, any
+identifier that is solely used there, can never be extracted.
+
+The fact that we can estimate all relevant identifiers for an article (see
+Section \ref{ir}), combined with some common assumptions about the relation,
+can be exploited to largely reduced set of candidates that need to be ranked.
+Please note that this reduction is essential for retrieving the correct
+relations. Otherwise almost any word would be ranked and the precision of the
+retrieval would drop significantly.
+
+The basic assumption of our approach is that if two entities take part in a
+relation, those two co-occur in the same sentence. Any other sentence is,
+therefor, irrelevant. Another assumption can be made about the lexical class
+of the desciption we want to rank. They are nouns or even noun phrases (e.g.
+\texttt{ `the effective spring constant k' } or \texttt{ `mass m of
+something'} ). Beside discarding all other words (according to their POS tag),
+noun phrases and terms within a Wikipedia hyperlink will be aggregated into
+one new candidate term. This is important, due to the fact that overall
+ranking will be greatly infuenced by the distance of candidate words. For all
+intends and purposes, it is not needed to keep noun phrases as single words.
+So they can be aggregated and be ranked as if they were a single word.
+
+\subsection{Numerical Statistics}
+
+Each description candidate is ranked with the weighted sum
+
 \begin{equation} \label{eq:rating}
@@
 +\beta{R}_{\sigma_\mathrm s}(n)
-+\gamma\mathrm{tf}(t,d)}{\alpha+\beta+\gamma} \mapsto [0;1].
++\gamma\mathrm{tf}(t,s)}{\alpha+\beta+\gamma} \mapsto [0;1].
 \end{equation}
-It depends on the distance $\Delta$ between identifier and definition term $t$, the distance $n$ between the display-style equation and the sentence that contains the term, and the identifier and the term frequency $\mathrm{tf}(t,d)$ in the current document $d$.
-The distance was normalized with $R_\sigma(\Delta)= \exp\left[-\frac{1}{2}\frac{\Delta^2-1}{\sigma^2}\right].$
-We assume that the probability to find a relation at $\Delta=1$ is maximal.  For example in the 
-text fragment \emph{the energy $E$, and the mass $m$,}  
-in order to determine the full width at half maximum of our distribution, we evaluated some articles manually and found $R_{\sigma_\mathrm d}(1)\approx 2 R_{\sigma_\mathrm d}(5)$ and thus 
-$\sigma_d=\sqrt\frac{12}{\ln 2}$.
-The probability to find a correct definition decays for 50\% within three sentences.
-Consequently  $\sigma_\mathrm s=2\left({\ln 2}\right)^{-\frac{1}{2}}$.
 
-The classic tf-idf \cite{Salton86} statistic reflects the importance of a term to a document.
-For our task, the inverse document frequency (idf) would assign high penalties to 
-frequent words like `energy', in comparison to 
-%\todo{I replaced `seldom words' with `words seldom seen'.  Is that ok?}
-words seldom seen like `Einstein', which are valid definitions for identifiers.
-As the influence of $\mathrm{tf}(t,d)$ on the overall ranking (\ref{eq:rating}) seems to be very high, we reduce the impact with the tuning parameters $\gamma=0.75$ and $\alpha = \beta = 1$.
+It depends on the distance $\Delta$ (= amount of word tokens) between
+identifier and the description term $t$, the sentence number $n$ counting
+(from the beginning of the article) all sentences containing the identifier,
+and the term frequency $\mathrm{tf}(t,s)$ in the set of sentences $s$. The
+distance was normalized with $R_\sigma(\Delta) = \exp\left[ -\frac{1}{2}
+\frac{\Delta^2-1}{\sigma^2}\right].$ We assume that the probability to find a
+relation at $\Delta=1$ is maximal. For example in the text fragment
+\texttt{`the energy E, and the mass m'}, in order to determine the full width
+at half maximum of our distribution, we evaluated some articles manually and
+found $R_{\sigma_\mathrm d}(1)\approx 2 R_{\sigma_\mathrm d}(5)$ and thus
+$\sigma_d=\sqrt\frac{12}{\ln 2}$. The probability to find a correct definition
+decays for 50\% within three sentences. Consequently $\sigma_\mathrm
+s=2\left({\ln 2}\right)^{-\frac{1}{2}}$.
+
+\paragraph{Robustness.}
+
+The classic tf-idf \cite{Salton86} statistic reflects the importance of a term
+to a document. For our task, the inverse document frequency (idf) would assign
+high penalties to frequent words like `length', in comparison to words seldom
+seen like `Hamiltonian', which are both valid definitions for identifiers. As the
+influence of $\mathrm{tf}(t,s)$ on the sensivity of the overall ranking
+(\ref{eq:rating}) seems to be very high, we reduce the impact with the tuning
+parameters $\gamma=0.1$ and $\alpha = \beta = 1$. Please note that this work
+is still in progress. For now the algorithm only takes the sentences into
+account, that were found in a single article. In the future, the retrieval
+process will take sets of closely related articles into account, trying
+leverage the problem that distributional properties will be volatile on term
+universes with very few members (e.g. term frequencies in a single sentence).
 
 \paragraph{Implementation.}
-We implemented the MLP processing system \cite{github} as Stratosphere data-flow using Java which allows for scalable execution, application of complex higher order functions and easy integration of third party tools like Stanford NLP and the Mylyn framework.
+
+We implemented the MLP processing system \cite{github} as Stratosphere data-
+flow using Java which allows for scalable execution, application of complex
+higher order functions and easy integration of third party tools like Stanford
+NLP and the Mylyn framework for mark-up parsing.
+
+\tikzset{actor/.style={
+        rectangle,
+        minimum size=6mm,
+        very thick,
+        draw=gray!50!black!50,
+        top color=white,
+        bottom color=gray!50!black!20
+    },
+    arrow/.style={
+        -latex, thick, shorten <=2pt,shorten >=2pt
+    }
+}
+\begin{figure}[H]
+	\begin{tikzpicture}[node distance=5mm and 8mm]
+		\node (Input) [align=center]{Wiki Dumps};
+		\node (DocumentParser) [actor, right=of Input, align=center] {\emph{Map}\\\textbf{Parser}};
+		\node (Candidates) [actor, right=of DocumentParser, align=center] {\emph{CoGroup}\\\textbf{Kernel}};
+		\node (Sentence) [actor, above=of DocumentParser, align=center] {\emph{Map}\\\textbf{Tagger}};
+		\node (Filter) [actor, right=of Candidates, align=center] {\emph{Reduce}\\\textbf{Filter}};
+		\node (Output) [right=of Filter,align=center]{Raw\ Candidates};
+		\draw[arrow] (Input)--(DocumentParser);
+		\draw[arrow] (DocumentParser)--(Sentence);
+		\draw[arrow] (DocumentParser)--(Candidates);
+		\draw[arrow] (Sentence.east)--(Candidates.north);
+		\draw[arrow] (Candidates)--(Filter);
+		\draw[arrow] (Filter)--(Output);
+	\end{tikzpicture}
+\caption{Data flow of the Stratosphere program}
+\end{figure}
+
 \section{Evaluation}
-We extracted 460 identifiers definition tuples in 550000 relations with a
-high score ($>0.8$).  The most common identifier definition tuples were 
-number $n$ (1709), time $t$ (1480), mass $M$ (1042), radius $r$ (752), 
-temperature $T$ (666), angle $\theta$ (639), group $G$ (635).
-Because we do not have any annotated test corpora,
-evaluation has to be performed by hand.
-Therefore, we took a sample of 30
-random documents and counted all matches.
-Most of the retrieved definitions were considered as relevant.
-Furthermore, for more than 50\% of the investigated identifiers, 
-the most probable meanings detected by our system were relevant.
-However, we neither consider identifiers with subscripts, nor identifiers that do not occur in the surrounding text.  
-%The resulting estimates for
-%\emph{recall} and \emph{precision} are 0.99 and 0.86 respectively.
+
+\subsubsection{Identifier Retrieval}
+\label{ir}
+
+Throughout our experiments we made some observations that had an impact on the
+accuracy of retrieving the correct set of indentifer. First of all, people
+tend to incorrectly use \emph{texvc} trying to create formulas. E.g.,
+\texttt{\textbackslash text\{log\}} is more often used than the correct
+operator \texttt{\textbackslash log}. Another problem is that sometimes people
+use indices as a form of `in field' annotation, like $T_{before}$ and
+$T_{after}$. The identifier $T$ is defined in the surrounding text, but
+neither $T_{before}$ nor $T_{after}$. There are more ambiguities. For example
+the superscripted 2 in $x^{2}$ and $\sigma^{2}$ can be interpreted as the
+power or as a part of the identifier. Another ambiguity is that the
+multiplication sign can be omitted, so that it is undecidable for a naive
+program whether $ab^{2}$ contains one or two identifiers.
+
+We took a very conservative approach and preprocessed all formulas.
+\texttt{\textbackslash text\{\}} blocks along with subscriptions containing
+more than a single character will be removed before analysis. Superscripts
+will also be ignored in terms of being a part of the identifier. Moreover, we
+created a comprehensive blacklist to improve the results further. Short words
+like `a' and `i', which are very common in the english language, could be
+easily matched by our processor in the sourrunding text if a formula contains
+one of those identifier, and therefor, will also be blacklisted. Additionally,
+we blacklist common mathematical operators, constants, and functions.
+
+Because of the fact, that we do not have any annotated test corpora,
+evaluation has to be performed by hand. Therefore, we took a sample of 30
+random articles and counted all matches. The resulting estimates for the
+identifier retrieval performance are: \emph{Recall: 0.99} and \emph{Precision:
+0.86}, which satisfy our information needs, as we are mostly interested in recall
+a this stage.
+
+\subsection{Description Retrieval}
+
+We ran the program on a dataset of about 20.000 articles, all containing
+\texttt{<math/>} tags, and retrieved about 550.000 candidate relations. The
+most common definition relations were:
+
+\begin{tabular}{| l | p{6.8cm} | l |}
+	\hline
+	Identifier & Descriptions & Count\\
+	\hline
+	$n$ & number & 1709 \\
+	$t$ & time & 1480 \\
+	$M$ & mass & 1042 \\
+	$r$ & radius & 752 \\
+	$T$ & temperature & 666 \\
+	$\theta$ & angle & 639 \\
+	$G$ & group & 635 \\
+	\hline
+\end{tabular}
+
+\paragraph{Observations.} We observed some poorly ranked relations. For
+example, in the fragment \texttt{`where $\phi$ ( $r_{i}$ ) is the electrostatic
+potential'} the distance
+$\Delta(\phi, \mathsf{electrostatic potential} ) = 5$. This is due
+to counting brackets and function arguments as words. Also wrongly tagged
+words like `Hamiltonian/JJ' lead to false negatives.
+
+\subsubsection{Comparative Evaluation}
+
+As there were no evaluation datasets available to us, we compared our
+statistical ranking to a simple pattern matching approach. We used the
+following patterns to discover description terms:
+
+\begin{tabular}{| p{9.3cm} |}
+	\hline
+	Pattern \\
+	\hline
+	\texttt{\emph{<description>} \emph{<identifier>}} \\
+	\texttt{\emph{<identifier>} is \emph{<description>}} \\
+	\texttt{\emph{<identifier>} is the \emph{<description>}} \\
+	\texttt{let \emph{<identifier>} be the \emph{<description>}} \\
+	\texttt{\emph{<description>} is|are denoted by \emph{<identifier>}} \\
+	\texttt{\emph{<identifier>} denotes */DT \emph{<identifier>}} \\
+	\hline
+\end{tabular}
+
+\paragraph{} Due to large amount of time needed to evaluate two approaches by hand on a
+multitude of articles, we build a dataset of definition relations (98
+relations in total), that can be retrieved from two representative articles
+(revision ids are included). This dataset is also available on the project
+repository.
+
+\begin{tabular}{| l | l | l | l |}
+	\hline
+	 & MLP-Ranking ($k=1$) & MLP-Ranking ($k=2$) & Pattern Matching \\
+	\hline
+	Precision &  0.872  &  0.915  &  0.911  \\
+	Recall    &  0.839  &  0.892  &  0.733  \\
+	\hline
+\end{tabular}
+
+Note: $k$ equals the amount of the top ranked candidate definitions.
+
+% \paragraph{} We than took the pattern matching approach as a baseline for our ranked
+% retrievals sets as these patterns cover most common definition relations we
+% found so far.
 
 \section{Further work}
-Our original intuition was to discover grammatical patterns 
-like \emph{let \texttt{variable} be \texttt{definition}} or 
-were \emph{\texttt{variable} (indicates/stands for/denotes) \texttt{definition}} based on the statistical findings.
-However, our impression is that this 
-would not lead to significant performance gain.
 
-The distance measure
-$R_{\sigma_d}$
-fails for the example of Fig.~\ref{fig:screenshot} since
-$\Delta(\mathsf{energy},E)=\Delta(\mathsf{mass},m)=2$.
-We plan to consider the grammatical structure of the sentence for an improved distance measure $\Delta^{+} : \Delta^+(\mathsf{energy},E)\ge\Delta^+(\mathsf{mass},m)$.
-Currently, we use the score $R$ to identify the most probably definition-identifier tuple on each page, even if it occurs multiple times on the page.
-For example, on the `Mass-energy equivalence' page, 21 sentences contain the combination of the identifier $E$ and the noun `energy'.
-A promising approach, is to use $R^\Sigma=\sum_{i=1}^n 2^{-i} R_i,$ where $R_i$ is a 
-sorted list, were $R_1$ is the highest ranked definition for that relation according to the current measure $R$.
-To improve the robustness of the \emph{term frequency} measure further, one
-might cluster sets of scientific documents, based on their specific fields of
-research.
+Our original intuition was to discover grammatical patterns like
+\texttt{`\emph{<identifier>} indicates/stands for/denotes
+\emph{<description>}'} based on the statistical findings. However, our
+impression is that this would not lead to significant performance gain.
+
+The distance measure $R_{\sigma_d}$ fails for the example of
+Fig.~\ref{fig:screenshot} since $\Delta(\mathsf{energy},E) =
+\Delta(\mathsf{mass},m) = 2$. Unfortunatly one cannot simply detect
+punctuation marks and introduce some kind of directed associativity (e.g.
+inflicting a penatly on the ranking if the relations spans over a comma). That
+would lead to whole classes of relation `types' (in terms of the grammatical
+structure) never being retrieved. We plan to mitigate this problem by taking
+more closely related scientific articles (based on their specific fields) into
+consideration and count the frequencies of the candidate relations. The
+intuition behind this is, that articles of the same scientific field will
+likely use the same definition for the identifiers.
+
+Currently, we use the ranking $R$ to identify the most probable description-
+identifier tuple on each article, even if it occurs multiple times on the page.
+For example, in the `Mass-energy equivalence' article, 21 sentences contain the
+combination of the identifier $E$ and the noun `energy'. A promising approach,
+is to use $R^\Sigma=\sum_{i=1}^n 2^{-i} R_i,$ where $R_i$ is a sorted list,
+were $R_1$ is the highest ranked definition for that relation according to the
+current measure $R$.
+
+Furthermore, a systematic approach for determining a wise choice of the
+ranking parameters should significantly improve the overall performance of our
+system.
 
 \section{Conclusion}
-Our experiments showed that combining a POS tagger with numerical statistics
-about the text surface, can lead to quality results.
-However, this approach is
-only applicable under certain conditions. 
-For identifiers which
-are seldom seen, our statistical approach tends to fail.
+
+Our experiments showed that selecting candidates according to their POS tags
+combined with numerical statistics about the text surface, can lead to quality
+results. However, this approach is only applicable under certain conditions.
+For identifiers which are seldom seen, our statistical approach tends to fail.
 In that situation, other methods, especially supervised ones, are used to
-bootstrap a language model.
-Unfortunately, they require a labeled test
-corpus to measure the performance of a classifier that could be trained with our generated data.
-Currently, we are planning to use the
-NTCIR-Math-10 Task, Math Understanding Subtask gold standard dataset \cite{overview}
-for a comparable evaluation.
-Furthermore, a systematic approach for determining a wise choice of the ranking parameters should significantly improve the overall performance of our system.
+bootstrap a language model. Unfortunately, they require a labeled test corpus
+to measure the performance of a classifier that could be trained with our
+generated data. Currently, we are planning to use the NTCIR-Math-10 Task, Math
+Understanding Subtask gold standard dataset \cite{overview} for a comparable
+evaluation.
+
+During this project we had the impression that one could discover `namespaces'
+(sets of documents, that use the same definitions for identifier) to aid in
+the retrieval procress. Robert Pagel currently works in his diploma thesis on
+this topic.
+
 \paragraph*{Acknowledgments.}
-Thanks to Howard Cohl for proof reading the paper and to Holmer Hemsen, the course instructor of the database project course at TU-Berlin in Fall 2012.
-The implementation and a first draft of this paper was completed in the duration of this course.
+
+Thanks to Howard Cohl for proof reading the paper and to Holmer Hemsen, the
+course instructor of the database project course at TU-Berlin in Fall 2012.
+The implementation and a first draft of this paper was completed in the
+duration of this course.
 

### 2014-06-15 paper.tex
current state of the article

@@
 \usetikzlibrary{arrows,backgrounds,positioning}
+\usepackage{todonotes}
 
@@
 \emph{Mathematical Language Processing} project aims to support that process
-and presents automated identifier definition discovery. Ordinary natural
-language processing approaches are trained on large, manually annotated text
-corpora. They fall short in reasonable quality of definition discovery for
-scientific texts. We use statistical methods to discover definitions in the
-surrounding text. The evaluation of our prototypical system, applied on the
+and presents automated identifier definition discovery. We compare two
+approaches. At first we use a simple identifier definition pattern matching
+approach. Second, we present the \emph{MLP} approach, that uses POS tag based
+distances as well as sentence positions to caculate indentifier definition
+probabilities. The evaluation of our prototypical system, applied on the
 Wikipedia text corpus, shows that our approach augments the user experience
@@
 
+\section{Pattern-based Definition Discovery}
+
+At first, we implemented a simple identifier definition extractor that is
+based on a set of static patterns. As it is an fairly robust approach und easy
+to implement, this will serve us as a good reference point in terms of
+performance. The following patterns are being used to discover description
+terms:
+
+\begin{table}
+\vspace{-5pt}
+	\begin{center}
+		\begin{tabular}{| p{9.3cm} |}
+			\hline
+			Pattern \\
+			\hline
+			\texttt{\emph{<description>} \emph{<identifier>}} \\
+			\texttt{\emph{<identifier>} is \emph{<description>}} \\
+			\texttt{\emph{<identifier>} is the \emph{<description>}} \\
+			\texttt{let \emph{<identifier>} be the \emph{<description>}} \\
+			\texttt{\emph{<description>} is|are denoted by \emph{<identifier>}} \\
+			\texttt{\emph{<identifier>} denotes */DT \emph{<description>}} \\
+			\hline
+		\end{tabular}
+	\end{center}
+\caption{Implemented static patterns}
+\vspace{-5pt}
+\end{table}
+
+Note that in the last pattern we used a \emph{part-of-speech} tag. Actually
+determiners not only contain articles but also quatifiers and distributives.
+To increase readability this list have been shortened to `*/DT'.
+
 \section{Statistical Definition Discovery}
@@
 Extracting relevant identifiers from the article rely on the assumption that
-the author will use \texttt{<math/>} tags for all formulae. Thus said, a
+the author will use \texttt{<math/>} tags for all formulae. This said, a
 formula that is written within a text cannot be recognized, and therefor, any
@@
 Section \ref{ir}), combined with some common assumptions about the relation,
-can be exploited to largely reduced set of candidates that need to be ranked.
+can be exploited to largely reduce the set of candidates that need to be ranked.
 Please note that this reduction is essential for retrieving the correct
@@
 
+The basic assumption of outr approach is that the two entities of a definition
+relation co-occur in the same sentence.
+
 The basic assumption of our approach is that if two entities take part in a
-relation, those two co-occur in the same sentence. Any other sentence is,
-therefor, irrelevant. Another assumption can be made about the lexical class
+relation, those two co-occur in the same sentence.
+Any other sentence is, therefor, irrelevant.
+
+Another assumption can be made about the lexical class
 of the desciption we want to rank. They are nouns or even noun phrases (e.g.
@@
 
-\subsubsection{Identifier Retrieval}
+\subsection{Identifier Retrieval}
 \label{ir}
@@
 created a comprehensive blacklist to improve the results further. Short words
-like `a' and `i', which are very common in the english language, could be
-easily matched by our processor in the sourrunding text if a formula contains
-one of those identifier, and therefor, will also be blacklisted. Additionally,
+like `a' and `i', which are very common in the English language, could be
+easily matched by our processor in the surrounding text if a formula contains
+one of those identifier, and therefore, will also be blacklisted. Additionally,
 we blacklist common mathematical operators, constants, and functions.
@@
 \subsection{Description Retrieval}
-
 We ran the program on a dataset of about 20.000 articles, all containing
@@
 
+\begin{table}[H]
+	\begin{center}
 	\begin{tabular}{| l | p{6.8cm} | l |}
@@
 	\end{tabular}
+	\end{center}
+\caption{Most common definition relations}
+\end{table}
 
@@
 potential'} the distance
-$\Delta(\phi, \mathsf{electrostatic potential} ) = 5$. This is due
+$\Delta(\phi, \mathsf{electrostatic\:potential} ) = 6$. This is due
 to counting brackets and function arguments as words. Also wrongly tagged
-words like `Hamiltonian/JJ' lead to false negatives.
+words like `Hamiltonian' as an adjective leads to false negatives. Latter
 
@@
 
-As there were no evaluation datasets available to us, we compared our
-statistical ranking to a simple pattern matching approach. We used the
-following patterns to discover description terms:
+Due to the lack of gold standard evaluation datasets to measure the
+performance of identifer definition extractors, we've created on on our own.
+This is a very time consuming job. At the moment the dataset only contains two
+large articles (revision ids are included) with around 100 identifier
+definitions. This dataset is also available on the project repository.
 
-\begin{tabular}{| p{9.3cm} |}
-	\hline
-	Pattern \\
-	\hline
-	\texttt{\emph{<description>} \emph{<identifier>}} \\
-	\texttt{\emph{<identifier>} is \emph{<description>}} \\
-	\texttt{\emph{<identifier>} is the \emph{<description>}} \\
-	\texttt{let \emph{<identifier>} be the \emph{<description>}} \\
-	\texttt{\emph{<description>} is|are denoted by \emph{<identifier>}} \\
-	\texttt{\emph{<identifier>} denotes */DT \emph{<identifier>}} \\
-	\hline
-\end{tabular}
+As in lots of articles, the ones in the evaluation dataset contain identifiers
+whose description cannot be retrieved. This is due to two reasons. First and
+foremost, the identifier found in a formula is never mentioned in the
+surrounding text, and therefor, no description can be extracted. Second, the
+identifier is somehow ambigious (see section~\ref{ir}) and has been dropped. Most
+notably, identifiers like $I_{xx}$ will be discarded because of an ambigious
+index that contains multiple letters.
 
-\paragraph{} Due to large amount of time needed to evaluate two approaches by hand on a
-multitude of articles, we build a dataset of definition relations (98
-relations in total), that can be retrieved from two representative articles
-(revision ids are included). This dataset is also available on the project
-repository.
+Unfortunately 32 out of 99 identifiers from our dataset fall into that category.
+We've decided to evaluate the performance of the remainder, as those 32 do not
+convey any conceptual flaws. From the users standpoint, the overall performance
+(in terms of recall) of such a system whould be rather annoying. As we are only
+interested in evaluating the performance of the \emph{MLP Ranking} algoritm itself,
+it is save to ignore those 32 identifers.
 
+\begin{table}[H]
+\vspace{-5pt}
+	\begin{center}
 		\begin{tabular}{| l | l | l | l |}
@@
 		\end{tabular}
+	\end{center}
+\caption{Evaluation results. Note: $k$ equals the amount of the top ranked candidate definitions.}
+\vspace{-20pt}
+\end{table}
 
-Note: $k$ equals the amount of the top ranked candidate definitions.
-
-% \paragraph{} We than took the pattern matching approach as a baseline for our ranked
-% retrievals sets as these patterns cover most common definition relations we
-% found so far.
+The results show that the unoptimized MLP approach keeps up with the
+performance of the simple pattern matcher. Furthermore, we observed that it is
+more robust in terms of recall, as it is less vulnerable to small changes in
+the sentence structure.
 
@@
 Fig.~\ref{fig:screenshot} since $\Delta(\mathsf{energy},E) =
-\Delta(\mathsf{mass},m) = 2$. Unfortunatly one cannot simply detect
+\Delta(\mathsf{mass},m) = 2$. Unfortunately, one cannot simply detect
 punctuation marks and introduce some kind of directed associativity (e.g.
-inflicting a penatly on the ranking if the relations spans over a comma). That
-would lead to whole classes of relation `types' (in terms of the grammatical
-structure) never being retrieved. We plan to mitigate this problem by taking
-more closely related scientific articles (based on their specific fields) into
-consideration and count the frequencies of the candidate relations. The
-intuition behind this is, that articles of the same scientific field will
-likely use the same definition for the identifiers.
+inflicting a penalty on the ranking if the candidate relations spans over a
+comma). That would lead to whole classes of relation `types' (in terms of the
+grammatical structure) never being retrieved. We plan to mitigate this problem
+by taking more closely related scientific articles (based on their specific
+fields) into consideration and count the frequencies of the candidate
+relations. The intuition behind this is, that articles of the same scientific
+field will likely use the same definition for the identifiers. Moreover,
+we hope to also mitigate the problem of `dangling' identifiers (those not
+mentioned in the article itself), as their may be description in related
+articles.
 

### 2014-06-15 paper.tex
Little changes in the abstract and introduction

@@
 Thus scientists must study the context to decode the meaning. The
-\emph{Mathematical Language Processing} project aims to support that process
-and presents automated identifier definition discovery. We compare two
-approaches. At first we use a simple identifier definition pattern matching
-approach. Second, we present the \emph{MLP} approach, that uses POS tag based
-distances as well as sentence positions to caculate indentifier definition
-probabilities. The evaluation of our prototypical system, applied on the
+\emph{Mathematical Language Processing} project aims to support that process.
+In this paper, we compare two approaches to discover identifier - definition tuples.
+At first we use a simple pattern matching
+approach. Second, we present the \emph{MLP} approach that uses POS tag based
+distances as well as sentence positions to calculate identifier - definition
+probabilities.
+The evaluation of our prototypical system, applied on the
 Wikipedia text corpus, shows that our approach augments the user experience
@@
 scientists. Often, they contain identifier whose meaning might be at first
-unknown or at least ambiguous to the reader (depending on their knownledge).
+unknown or at least ambiguous to the reader (depending on their knowledge).
 Therefore, one usually needs to study the surrounding text to find the
@@
 especially profit from a system that helps them to understand formulae more
-quickly.
+quickly. In the long term the extracted identifier definition tuples contribute
+to an increase of machine readability of scientific publications. This builds a 
+foundation for added value services such as search, clustering and improved
+accessibility.
 

### 2014-06-17 paper.tex
Sec 2 is not unerstandable

@@
 
-In natural language, words and phrases themselves imply the semantics. In
-contrast, the meaning of identifiers in mathematical formulae is undefined.
-Thus scientists must study the context to decode the meaning. The
-\emph{Mathematical Language Processing} project aims to support that process.
-In this paper, we compare two approaches to discover identifier - definition tuples.
-At first we use a simple pattern matching
-approach. Second, we present the \emph{MLP} approach that uses POS tag based
-distances as well as sentence positions to calculate identifier - definition
-probabilities.
-The evaluation of our prototypical system, applied on the
-Wikipedia text corpus, shows that our approach augments the user experience
-substantially. While hovering the identifiers in the formula, tool-tips with
-the most probable definitions occur. Tests with random samples show that the
-displayed definitions provide a good match with the actual meaning of the
-identifiers.
+In natural language, words and phrases themselves imply the semantics. In contrast, the meaning of identifiers in mathematical formulae is undefined.
+Thus scientists must study the context to decode the meaning.
+The \emph{Mathematical Language Processing} (MLP) project aims to support that process.
+In this paper, we compare two approaches to discover identifier-definition tuples.
+At first we use a simple pattern matching approach.
+Second, we present the \emph{MLP} approach that uses Part-of-speech (POS) tag based distances as well as sentence positions to calculate indentifier-definition probabilities.
+The evaluation of our prototypical system, applied on the Wikipedia text corpus, shows that our approach augments the user experience substantially.
+While hovering the identifiers in the formula, tool-tips with the most probable definitions occur.
+Tests with random samples show that the displayed definitions provide a good match with the actual meaning of the identifiers.
 
@@
 
-To build such a system, a labeled text corpus that annotates identifiers and
+To build such a system, a labelled text corpus that annotates identifiers and
 their definition is desirable. At the project start, such a corpus was not
-available. We use simple statistical assumptions about co-occurrences between
-identifiers and their potential description to extract the meaning of the
-identifiers from their context. In combination with the Stratosphere system
+available. 
+We assume that the probability to find correct identifier definition tuples is
+POS tagged based distance metrics.
+Our motivation for that is readers usually extract the identifier definition
+from the context that is given by the surrounding text.
+In combination with the Stratosphere system
 that is capable of scalable data processing,  we rapidly generate a large
 amount of definition relation candidates with only minimal implementation
-overhead to ensure the scalability of our algorithm.
+overhead for the parallelization.
 
@@
 most articles make use of \texttt{<math/>} tags (\emph{texvc} as an input
-language) for formulae. It is easy to detect and the extraction of identifiers
-is well tried. Second, the text is already annotated with mark-up.
+language) for formulae. Those tags are easy to detect and the extraction identifiers
+in its experimental MathML output format is well tried. 
+Second, the text is already annotated with mark-up.
 Particularly, hyperlinks to other articles within the Wikipedia are of
@@
 At first, we implemented a simple identifier definition extractor that is
-based on a set of static patterns. As it is an fairly robust approach und easy
+based on a set of static patterns. As it is an fairly robust approach and easy
 to implement, this will serve us as a good reference point in terms of
@@
 Note that in the last pattern we used a \emph{part-of-speech} tag. Actually
-determiners not only contain articles but also quatifiers and distributives.
+determiners not only contain articles but also quantifiers and distributives.
 To increase readability this list have been shortened to `*/DT'.
+\todo[inline]{describe the process: extraction of identifiers, how to didentify the dscription? Looking for space or using pos tags?}
 

### 2014-06-17 paper.tex
Corrections and comments from Howard

--- 2014-06-15 paper.tex
@@
 
-In natural language, words and phrases themselves imply the semantics. In
-contrast, the meaning of identifiers in mathematical formulae is undefined.
-Thus scientists must study the context to decode the meaning. The
-\emph{Mathematical Language Processing} project aims to support that process.
-In this paper, we compare two approaches to discover identifier - definition tuples.
-At first we use a simple pattern matching
-approach. Second, we present the \emph{MLP} approach that uses POS tag based
-distances as well as sentence positions to calculate identifier - definition
-probabilities.
-The evaluation of our prototypical system, applied on the
-Wikipedia text corpus, shows that our approach augments the user experience
-substantially. While hovering the identifiers in the formula, tool-tips with
-the most probable definitions occur. Tests with random samples show that the
-displayed definitions provide a good match with the actual meaning of the
-identifiers.
+In natural language, words and phrases themselves imply the semantics. In contrast, the meaning of identifiers in mathematical formulae is undefined.
+Thus scientists must study the context to decode the meaning.
+The Mathematical Language Processing (MLP) project aims to support that process.
+In this paper, we compare two approaches to discover identifier-definition tuples.
+At first we use a simple pattern matching approach.
+Second, we present the
+MLP approach that uses part-of-speech  tag based distances as well as sentence positions to calculate identifier-definition probabilities.
+The evaluation of our prototypical system, applied on the Wikipedia text corpus, shows that our approach augments the user experience substantially.
+While hovering the identifiers in the formula, tool-tips with the most probable definitions occur.
+Tests with random samples show that the displayed definitions provide a good match with the actual meaning of the identifiers.
 
@@
 
-Mathematical formulae are a viable source of information for a wide range of
-scientists. Often, they contain identifier whose meaning might be at first
+Mathematical formulae are viable sources of information for a wide range of
+scientists. Often, they contain identifiers whose meaning might be at first
 unknown or at least ambiguous to the reader (depending on their knowledge).
@@
 relevant definition. An automatic information retrieval system can be used to
-reduce the reader's effort by showing the most relevant definition relation
-found to the reader. Students and scientists of other disciplines would
+reduce the reader's effort by displaying the most relevant definition relation
+found, to the reader. Students and scientists of other disciplines would
 especially profit from a system that helps them to understand formulae more
-quickly. In the long term the extracted identifier definition tuples contribute
-to an increase of machine readability of scientific publications. This builds a 
+quickly. In the long term, the extracted identifier definition tuples contribute
+to an increased machine readability of scientific publications. This builds a 
 foundation for added value services such as search, clustering and improved
@@
 their definition is desirable. At the project start, such a corpus was not
-available. We use simple statistical assumptions about co-occurrences between
-identifiers and their potential description to extract the meaning of the
-identifiers from their context. In combination with the Stratosphere system
+available. 
+We assume that the probability to find correct identifier definition tuples is
+part-of-speech (POS) tagged based distance metrics.
+Our motivation for our choice is that readers usually extract identifier 
+definitions from context that is given by the surrounding text.
+In combination with the Stratosphere system \todo{[HSC: NEED REFERENCE]}
 that is capable of scalable data processing,  we rapidly generate a large
 amount of definition relation candidates with only minimal implementation
-overhead to ensure the scalability of our algorithm.
+overhead for the parallelization.
 
@@
 most articles make use of \texttt{<math/>} tags (\emph{texvc} as an input
-language) for formulae. It is easy to detect and the extraction of identifiers
-is well tried. Second, the text is already annotated with mark-up.
-Particularly, hyperlinks to other articles within the Wikipedia are of
+language) for formulae.  The identification of these tags are easy to find, 
+and from the MathML output, it is easy to extract the identifiers.
+Second, the text is already annotated with mark-up.
+Particularly, hyperlinks to other articles within Wikipedia are of
 interest as they typically wrap around any number of words and indicate that
-these in combination are relevant in the given context (respectively
-sentence).
+these in combination are relevant in the given context or (respectively) sentence.
 
@@
 compute with tens of thousands of articles. Especially when using text
-annotators (e.g. \emph{POS} tagger \cite{Rathna96}), like Stanford's NLP
+annotators (e.g., \emph{POS} tagger \cite{Rathna96}), like Stanford's NLP
 framework, one can make use of a parallel processing system to speed up
@@
 At first, we implemented a simple identifier definition extractor that is
-based on a set of static patterns. As it is an fairly robust approach und easy
-to implement, this will serve us as a good reference point in terms of
+based on a set of static patterns. As this is a fairly robust approach and easy
+to implement, it serves as a good reference point in terms of
 performance. The following patterns are being used to discover description
-terms:
+terms: \todo{[HSC: here you should refer to the table]}
 
@@
 
-Note that in the last pattern we used a \emph{part-of-speech} tag. Actually
-determiners not only contain articles but also quatifiers and distributives.
-To increase readability this list have been shortened to `*/DT'.
+Note that in the last pattern we used a POS tag.  Determiners not only contain 
+articles, but also quantifiers and distributives.
+To increase readability, this list have been shortened to `*/DT'. 
+\todo[inline]{[HSC: Explain */DT in detail.] describe the process: extraction of identifiers, 
+how to identify the description? Looking for space or using POS tags?}
 
@@
 We detect relations between identifiers and their description in two steps.
-First, we extract the identifiers from the formulae found in an article and
+First, we extract the identifiers from the formulae found in an article, and
 second we determine their description from the surrounding text.
@@
 the author will use \texttt{<math/>} tags for all formulae. This said, a
-formula that is written within a text cannot be recognized, and therefor, any
-identifier that is solely used there, can never be extracted.
+formula that is written within the text cannot be recognized, and therefore
+can never be extracted.
 
@@
 
-The basic assumption of outr approach is that the two entities of a definition
+\todo{HSC: The following two sentences need to be merged and explained further.
+Also, paragraphs need to have more than one sentence. Fix this.}
+The basic assumption of our approach is that the two entities of a definition
 relation co-occur in the same sentence.
@@
 Another assumption can be made about the lexical class
-of the desciption we want to rank. They are nouns or even noun phrases (e.g.
-\texttt{ `the effective spring constant k' } or \texttt{ `mass m of
-something'} ). Beside discarding all other words (according to their POS tag),
-noun phrases and terms within a Wikipedia hyperlink will be aggregated into
-one new candidate term. This is important, due to the fact that overall
-ranking will be greatly infuenced by the distance of candidate words. For all
-intends and purposes, it is not needed to keep noun phrases as single words.
-So they can be aggregated and be ranked as if they were a single word.
+of the description we want to rank. They are nouns or even noun phrases 
+(e.g.,
+\texttt{`the effective spring constant k'} or \texttt{`mass m of
+something'}). We discarded all other words (according to their POS tag)
+except noun phrases and Wikipedia hyperlinks.  These are candidates for a 
+definition.  This is important, due to the fact that overall
+ranking will be greatly influenced by the distance of candidate words. For all
+intents and purposes, it is not necessary to keep noun phrases as single words.
+They can therefore be aggregated and be ranked as if they were a single word.
 
@@
 +\beta{R}_{\sigma_\mathrm s}(n)
-+\gamma\mathrm{tf}(t,s)}{\alpha+\beta+\gamma} \mapsto [0;1].
++\gamma\,\mathrm{tf}(t,s)}{\alpha+\beta+\gamma} \mapsto [0,1].
 \end{equation}
 
-It depends on the distance $\Delta$ (= amount of word tokens) between
+The weighted sum depends on the distance $\Delta$ (amount of word tokens) between
 identifier and the description term $t$, the sentence number $n$ counting
@@
 $\sigma_d=\sqrt\frac{12}{\ln 2}$. The probability to find a correct definition
-decays for 50\% within three sentences. Consequently $\sigma_\mathrm
+decays to 50\% within three sentences. Consequently $\sigma_\mathrm
 s=2\left({\ln 2}\right)^{-\frac{1}{2}}$.
@@
 The classic tf-idf \cite{Salton86} statistic reflects the importance of a term
-to a document. For our task, the inverse document frequency (idf) would assign
-high penalties to frequent words like `length', in comparison to words seldom
-seen like `Hamiltonian', which are both valid definitions for identifiers. As the
-influence of $\mathrm{tf}(t,s)$ on the sensivity of the overall ranking
+to a document. For our task, the inverse document frequency (idf) assigns
+high penalties to frequent words like `length', as opposed to words seldom
+seen such as `Hamiltonian'. These are both valid definitions for identifiers. 
+As the influence of $\mathrm{tf}(t,s)$ on the sensitivity of the overall ranking
 (\ref{eq:rating}) seems to be very high, we reduce the impact with the tuning
-parameters $\gamma=0.1$ and $\alpha = \beta = 1$. Please note that this work
-is still in progress. For now the algorithm only takes the sentences into
-account, that were found in a single article. In the future, the retrieval
-process will take sets of closely related articles into account, trying
-leverage the problem that distributional properties will be volatile on term
-universes with very few members (e.g. term frequencies in a single sentence).
+parameters $\gamma=0.1$ and $\alpha = \beta = 1$. Please note that the algorithm 
+currently only takes sentences into account which were found in a single article. 
+In the future, the retrieval process will examine sets of closely related articles.
+This will leverage the problem that distributional properties will be volatile on 
+term universes with very few members (e.g., term frequencies in a single sentence).
 
@@
 
-We implemented the MLP processing system \cite{github} as Stratosphere data-
-flow using Java which allows for scalable execution, application of complex
-higher order functions and easy integration of third party tools like Stanford
-NLP and the Mylyn framework for mark-up parsing.
+We implemented the MLP processing system \cite{github} as Stratosphere 
+data-flow using Java which allows for scalable execution, application of complex
+higher order functions, and easy integration of third party tools such as
+Stanford NLP and the Mylyn framework for mark-up parsing.
 
@@
 
-Throughout our experiments we made some observations that had an impact on the
-accuracy of retrieving the correct set of indentifer. First of all, people
-tend to incorrectly use \emph{texvc} trying to create formulas. E.g.,
+Throughout our experiments, we made some observations that had an impact on the
+accuracy of retrieving the correct set of identifiers. First of all, people
+tend to incorrectly use \emph{texvc} when trying to create formulas. For example, 
 \texttt{\textbackslash text\{log\}} is more often used than the correct
@@
 We took a very conservative approach and preprocessed all formulas.
-\texttt{\textbackslash text\{\}} blocks along with subscriptions containing
+The \TeX~command \texttt{\textbackslash text\{\}} blocks along with subscriptions containing
 more than a single character will be removed before analysis. Superscripts
@@
 created a comprehensive blacklist to improve the results further. Short words
-like `a' and `i', which are very common in the English language, could be
+like `a', `A', and `I', which are very common in the English language, could be
 easily matched by our processor in the surrounding text if a formula contains
-one of those identifier, and therefore, will also be blacklisted. Additionally,
+one of those identifiers, and therefore, will also be blacklisted. Additionally,
 we blacklist common mathematical operators, constants, and functions.
 
-Because of the fact, that we do not have any annotated test corpora,
-evaluation has to be performed by hand. Therefore, we took a sample of 30
+Because we do not have any annotated test corpora, evaluation has to be performed by hand. 
+Therefore, we took a sample of 30
 random articles and counted all matches. The resulting estimates for the
@@
 0.86}, which satisfy our information needs, as we are mostly interested in recall
-a this stage.
+at this stage.
 
 \subsection{Description Retrieval}
-We ran the program on a dataset of about 20.000 articles, all containing
-\texttt{<math/>} tags, and retrieved about 550.000 candidate relations. The
+We ran our program on a dataset of about 20,000 articles, all containing
+\texttt{<math/>} tags, and retrieved about 550,000 candidate relations. The
 most common definition relations were:
@@
 \paragraph{Observations.} We observed some poorly ranked relations. For
-example, in the fragment \texttt{`where $\phi$ ( $r_{i}$ ) is the electrostatic
-potential'} the distance
-$\Delta(\phi, \mathsf{electrostatic\:potential} ) = 6$. This is due
+example, in the fragment \texttt{`where $\phi$( $r_{i}$ ) is the electrostatic
+potential'}, the distance $\Delta(\phi, \mathsf{electrostatic\:potential} ) = 6$. This is due
 to counting brackets and function arguments as words. Also wrongly tagged
 words like `Hamiltonian' as an adjective leads to false negatives. Latter
+\todo{Latter ........ ???}
 
@@
 
-Due to the lack of gold standard evaluation datasets to measure the
-performance of identifer definition extractors, we've created on on our own.
-This is a very time consuming job. At the moment the dataset only contains two
+There is no gold standard evaluation datasets to measure the
+performance of identifier definition extractors. Thus created on on our own dataset.
+This is a very time consuming job. At the moment, the dataset only contains two
 large articles (revision ids are included) with around 100 identifier
@@
 
-As in lots of articles, the ones in the evaluation dataset contain identifiers
+As in many articles, those in the evaluation dataset contain identifiers
 whose description cannot be retrieved. This is due to two reasons. First and
 foremost, the identifier found in a formula is never mentioned in the
-surrounding text, and therefor, no description can be extracted. Second, the
-identifier is somehow ambigious (see section~\ref{ir}) and has been dropped. Most
-notably, identifiers like $I_{xx}$ will be discarded because of an ambigious
+surrounding text, and therefore, no description can be extracted. Second, the
+identifier is somehow ambiguous (see Section~\ref{ir}) and has been dropped. Most
+notably, identifiers like $I_{xx}$ will be discarded because of an ambiguous
 index that contains multiple letters.
@@
 convey any conceptual flaws. From the users standpoint, the overall performance
-(in terms of recall) of such a system whould be rather annoying. As we are only
-interested in evaluating the performance of the \emph{MLP Ranking} algoritm itself,
-it is save to ignore those 32 identifers.
+(in terms of recall) of such a system would be rather annoying. As we are only
+interested in evaluating the performance of the \emph{MLP Ranking} algorithm itself,
+it is safe to ignore those 32 identifiers.
 
@@
 
-The results show that the unoptimized MLP approach keeps up with the
+Our results show that the unoptimized MLP approach keeps up with the
 performance of the simple pattern matcher. Furthermore, we observed that it is
@@
 \Delta(\mathsf{mass},m) = 2$. Unfortunately, one cannot simply detect
-punctuation marks and introduce some kind of directed associativity (e.g.
-inflicting a penalty on the ranking if the candidate relations spans over a
-comma). That would lead to whole classes of relation `types' (in terms of the
+punctuation marks and introduce some kind of directed associativity 
+(e.g., inflicting a penalty on the ranking if the candidate relations spans over a
+comma). This leads to whole classes of relation `types' (in terms of the
 grammatical structure) never being retrieved. We plan to mitigate this problem
@@
 field will likely use the same definition for the identifiers. Moreover,
-we hope to also mitigate the problem of `dangling' identifiers (those not
-mentioned in the article itself), as their may be description in related
+we hope also to resolve the problem of `dangling' identifiers (those not
+mentioned in the article itself), as there may be a description in related
 articles.
@@
 combination of the identifier $E$ and the noun `energy'. A promising approach,
-is to use $R^\Sigma=\sum_{i=1}^n 2^{-i} R_i,$ where $R_i$ is a sorted list,
-were $R_1$ is the highest ranked definition for that relation according to the
-current measure $R$.
-
-Furthermore, a systematic approach for determining a wise choice of the
+is to use $R^\Sigma=\sum_{i=1}^n 2^{-i} R_i,$ where $R_i$ is a sorted list.
+Here, $R_1$ is the highest ranked definition for that relation according to the
+current measure $R$.  A systematic approach for determining a wise choice of the
 ranking parameters should significantly improve the overall performance of our
@@
 (sets of documents, that use the same definitions for identifier) to aid in
-the retrieval procress. Robert Pagel currently works in his diploma thesis on
-this topic.
+the retrieval process. Robert Pagel is currently working on this topic
+for his diploma thesis.
 
@@
 
-Thanks to Howard Cohl for proof reading the paper and to Holmer Hemsen, the
+Thanks to Howard Cohl for proofreading the paper and to Holmer Hemsen, the
 course instructor of the database project course at TU-Berlin in Fall 2012.
@@
 duration of this course.
+
+\todo{Fix two (?) dates in reference [3].}
 

### 2014-06-19 paper.tex
…

@@
 
-In natural language, words and phrases themselves imply the semantics. In contrast, the meaning of identifiers in mathematical formulae is undefined.
-Thus scientists must study the context to decode the meaning.
-The Mathematical Language Processing (MLP) project aims to support that process.
-In this paper, we compare two approaches to discover identifier-definition tuples.
-At first we use a simple pattern matching approach.
-Second, we present the
-MLP approach that uses part-of-speech  tag based distances as well as sentence positions to calculate identifier-definition probabilities.
-The evaluation of our prototypical system, applied on the Wikipedia text corpus, shows that our approach augments the user experience substantially.
-While hovering the identifiers in the formula, tool-tips with the most probable definitions occur.
-Tests with random samples show that the displayed definitions provide a good match with the actual meaning of the identifiers.
+In natural language, words and phrases themselves imply the semantics. In
+contrast, the meaning of identifiers in mathematical formulae is undefined.
+Thus scientists must study the context to decode the meaning. The Mathematical
+Language Processing (MLP) project aims to support that process. In this
+paper, we compare two approaches to discover identifier-definition tuples. At
+first we use a simple pattern matching approach. Second, we present the MLP
+approach that uses part-of-speech tag based distances as well as sentence
+positions to calculate identifier-definition probabilities. The evaluation of
+our prototypical system, applied on the Wikipedia text corpus, shows that our
+approach augments the user experience substantially. While hovering the
+identifiers in the formula, tool-tips with the most probable definitions
+occur. Tests with random samples show that the displayed definitions provide a
+good match with the actual meaning of the identifiers.
 
@@
 available.
-We assume that the probability to find correct identifier definition tuples is
-part-of-speech (POS) tagged based distance metrics.
-Our motivation for our choice is that readers usually extract identifier 
-definitions from context that is given by the surrounding text.
-In combination with the Stratosphere system \todo{[HSC: NEED REFERENCE]}
-that is capable of scalable data processing,  we rapidly generate a large
-amount of definition relation candidates with only minimal implementation
-overhead for the parallelization.
+We calculate the probablities for correct identifier definition tuples based
+on part-of-speech (POS) tagged based distance metrics.
+\todo{rewrite sentences to make it process more clear}
+ The motivation for our
+choice is that readers usually extract identifier definitions from context
+that is given by the surrounding text. In combination with the Stratosphere
+system \cite{Alexandrov2010} that is capable of scalable data processing, we
+rapidly generate a large amount of definition relation candidates with only
+minimal implementation overhead for the parallelization.
 
-We chose the Wikipedia text corpus as a target because of two facts. First,
+As the target text corpus we chose the Wikipedia because of two facts. First,
 most articles make use of \texttt{<math/>} tags (\emph{texvc} as an input
@@
 to implement, it serves as a good reference point in terms of
-performance. The following patterns are being used to discover description
-terms: \todo{[HSC: here you should refer to the table]}
+performance. It simply iterates through the text, trying to find word groups,
+that are matched by a pattern. The patterns being used to discover description
+terms are depicted in table \ref{tpatterns}. Due to the fact that we already
+tokenized an annotated the articles in a previous step in the MLP system, we
+can make use of POS tags here as well.
+
+Note, determiners not only contain articles, but also quantifiers and
+distributives. The last pattern in table \ref{tpatterns} contains `*/DT'. This
+is a shorthand for every word, that has the POS tag `DT' (determiner), will be
+matched at this position. Otherwise this pattern whould be rather large, as it
+needs to contain every possible determiner. \emph{identifer} as well as
+\emph{description} are placholders, that mark the positions of the entities
+from a possible defintion relation.
 
@@
 	\end{center}
-\caption{Implemented static patterns}
+\caption{\label{tpatterns}Implemented static patterns}
 \vspace{-5pt}
 \end{table}
-
-Note that in the last pattern we used a POS tag.  Determiners not only contain 
-articles, but also quantifiers and distributives.
-To increase readability, this list have been shortened to `*/DT'. 
-\todo[inline]{[HSC: Explain */DT in detail.] describe the process: extraction of identifiers, 
-how to identify the description? Looking for space or using POS tags?}
 
@@
 formula that is written within the text cannot be recognized, and therefore
-can never be extracted.
+can never be extracted by our system.
 
@@
 Section \ref{ir}), combined with some common assumptions about the relation,
-can be exploited to largely reduce the set of candidates that need to be ranked.
-Please note that this reduction is essential for retrieving the correct
-relations. Otherwise almost any word would be ranked and the precision of the
-retrieval would drop significantly.
+can be exploited to largely reduce the set of candidates that need to be
+ranked. Please note that this reduction is essential for retrieving the
+correct relations for our approach. Otherwise almost any word would be ranked
+and the precision of the retrieval would drop significantly.
 
-\todo{HSC: The following two sentences need to be merged and explained further.
-Also, paragraphs need to have more than one sentence. Fix this.}
 The basic assumption of our approach is that the two entities of a definition
-relation co-occur in the same sentence.
+relation co-occur in the same sentence. In other words, if we want to retrieve
+the description for an identifier, in the set of sentences that contain the
+identifier the description should be found. Thus said, any other sentences can
+be ignored. Furthermore, we assume that it is more likely to find the
+description in first sentences than in the latter. This is based on the idea
+that authors introduce the meaning of an identifer and than subsequent use it,
+without necessarily repeating the meaning.
 
-The basic assumption of our approach is that if two entities take part in a
-relation, those two co-occur in the same sentence.
-Any other sentence is, therefor, irrelevant.
-
-Another assumption can be made about the lexical class
-of the description we want to rank. They are nouns or even noun phrases 
-(e.g.,
-\texttt{`the effective spring constant k'} or \texttt{`mass m of
-something'}). We discarded all other words (according to their POS tag)
-except noun phrases and Wikipedia hyperlinks.  These are candidates for a 
-definition.  This is important, due to the fact that overall
-ranking will be greatly influenced by the distance of candidate words. For all
-intents and purposes, it is not necessary to keep noun phrases as single words.
-They can therefore be aggregated and be ranked as if they were a single word.
+Another assumption can be made about the lexical class of the description we
+want to rank. They are nouns or even noun phrases (e.g., \texttt{`the
+effective spring constant k'} or \texttt{`mass m of something'}). We discard
+all other words (according to their POS tag) except noun phrases and Wikipedia
+hyperlinks. These are the candidates for a definition relation. Noun phrases
+and hyperlinks may consist of multiple words. For all intents and purposes, it
+is not necessary to threat noun phrases and hyperlinks as a set of words, and
+therefore, will be threated subsequently as if they were one. This is
+important, due to the fact that the overall ranking will be greatly influenced
+by the distance of candidates to the position of the identifer.
 
@@
 (\ref{eq:rating}) seems to be very high, we reduce the impact with the tuning
-parameters $\gamma=0.1$ and $\alpha = \beta = 1$. Please note that the algorithm 
+parameters $\gamma=0.1$ and remain $\alpha = \beta = 1$. Please note that the algorithm
 currently only takes sentences into account which were found in a single article.
-In the future, the retrieval process will examine sets of closely related articles.
+In the future, the MLP system will examine sets of closely related articles.
 This will leverage the problem that distributional properties will be volatile on
@@
 
-Because we do not have any annotated test corpora, evaluation has to be performed by hand. 
-Therefore, we took a sample of 30
-random articles and counted all matches. The resulting estimates for the
-identifier retrieval performance are: \emph{Recall: 0.99} and \emph{Precision:
-0.86}, which satisfy our information needs, as we are mostly interested in recall
-at this stage.
+Because we do not have any annotated test corpora, evaluation has to be
+performed by hand. Therefore, we took a sample of 30 random articles and
+counted all matches. The resulting estimates for the identifier retrieval
+performance are: \emph{Recall: 0.99} and \emph{Precision: 0.86}, which satisfy
+our information needs, as we are mostly interested in recall at this stage.
 
@@
 \texttt{<math/>} tags, and retrieved about 550,000 candidate relations. The
-most common definition relations were:
+most common definition relations are listed in table \ref{tcommon}.
 
@@
 	\end{center}
-\caption{Most common definition relations}
+\caption{\label{tcommon}Most common definition relations}
 \end{table}
@@
 to counting brackets and function arguments as words. Also wrongly tagged
-words like `Hamiltonian' as an adjective leads to false negatives. Latter
-\todo{Latter ........ ???}
+words like `Hamiltonian' as an adjective leads to false negatives.
 
@@
 
-There is no gold standard evaluation datasets to measure the
-performance of identifier definition extractors. Thus created on on our own dataset.
-This is a very time consuming job. At the moment, the dataset only contains two
+There is no gold standard evaluation datasets to measure the performance of
+identifier definition extractors. Thus we created on on our own dataset. This
+is a very time consuming job. At the moment, the dataset only contains two
 large articles (revision ids are included) with around 100 identifier
@@
 field will likely use the same definition for the identifiers. Moreover,
-we hope also to resolve the problem of `dangling' identifiers (those not
-mentioned in the article itself), as there may be a description in related
+we also hope to resolve the problem of `dangling' identifiers (those not
+mentioned in the article itself), as they might be described in related
 articles.
@@
 For identifiers which are seldom seen, our statistical approach tends to fail.
-In that situation, other methods, especially supervised ones, are used to
-bootstrap a language model. Unfortunately, they require a labeled test corpus
-to measure the performance of a classifier that could be trained with our
-generated data. Currently, we are planning to use the NTCIR-Math-10 Task, Math
-Understanding Subtask gold standard dataset \cite{overview} for a comparable
-evaluation.
+In that situation, other methods, especially supervised ones, are preferred.
+Unfortunately, many of them require a labeled test corpus to measure the
+performance of a classifier that could be trained with our generated data.
+Currently, we are planning to use the NTCIR-Math-10 Task, Math Understanding
+Subtask gold standard dataset \cite{overview} for a comparable evaluation.
 
@@
 
-\todo{Fix two (?) dates in reference [3].}
-
 \begingroup

### 2014-06-20 paper.tex
fix final todo

@@
 \usetikzlibrary{arrows,backgrounds,positioning}
-\usepackage{todonotes}
 
@@
 available.
-We calculate the probablities for correct identifier definition tuples based
+Consequently we had to start manual investigation of individual articles.
+Our first observation was many identifier definitions use a fixed string pattern
+to explain the definition to the reader. Furhtermore most
+definitions are usually very close to the occurence of the identifier.
+Thus we calculate the probablities for correct identifier definition tuples based
 on part-of-speech (POS) tagged based distance metrics.
-\todo{rewrite sentences to make it process more clear}
- The motivation for our
-choice is that readers usually extract identifier definitions from context
+This correlates to the experience that readers usually extract identifier definitions from context
 that is given by the surrounding text. In combination with the Stratosphere
@@
 example, in the fragment \texttt{`where $\phi$( $r_{i}$ ) is the electrostatic
-potential'}, the distance $\Delta(\phi, \mathsf{electrostatic\:potential} ) = 6$. This is due
+potential'}, the distance is %\newline
+$\Delta(\phi, \mathsf{electrostatic\:potential} ) = 6$. This is due
 to counting brackets and function arguments as words. Also wrongly tagged

### 2014-06-30 paper.tex
typo & some quirky sentences rewritten

@@
 reduce the reader's effort by displaying the most relevant definition relation
-found, to the reader. Students and scientists of other disciplines would
+found to the reader. Students and scientists of other disciplines would
 especially profit from a system that helps them to understand formulae more
@@
 
-To build such a system, a labeled text corpus that annotates identifiers and
+To build such a system, a labelled text corpus that annotates identifiers and
 their definition is desirable. At the project start, such a corpus was not
-available.
-Consequently we had to start manual investigation of individual articles.
-Our first observation was many identifier definitions use a fixed string pattern
-to explain the definition to the reader. Furhtermore most
-definitions are usually very close to the occurence of the identifier.
-Thus we calculate the probablities for correct identifier definition tuples based
-on part-of-speech (POS) tagged based distance metrics.
-This correlates to the experience that readers usually extract identifier definitions from context
-that is given by the surrounding text. In combination with the Stratosphere
-system \cite{Alexandrov2010} that is capable of scalable data processing, we
-rapidly generate a large amount of definition relation candidates with only
-minimal implementation overhead for the parallelization.
+available. Consequently we had to start manual investigation of individual
+articles. Our first observation was that many identifier definitions use a
+fixed string pattern to explain the definition to the reader. Furthermore,
+most definitions usually appear very close to the related identifier in the
+sentences. Thus, we calculate the probabilities for correct identifier
+definition tuples based on distance metrics for certain part-of-speech (POS)
+tagged words. This correlates to the experience that readers usually extract
+identifier definitions from context that is given by the surrounding text.
 
-As the target text corpus we chose the Wikipedia because of two facts. First,
+We chose the Wikipedia as the target text corpus because of two facts. First,
 most articles make use of \texttt{<math/>} tags (\emph{texvc} as an input
-language) for formulae. The identification of these tags are easy to find,
-and from the MathML output, it is easy to extract the identifiers.
-Second, the text is already annotated with mark-up.
-Particularly, hyperlinks to other articles within Wikipedia are of
-interest as they typically wrap around any number of words and indicate that
-these in combination are relevant in the given context or (respectively) sentence.
+language) for formulae. The identification of \texttt{<math/>} tags is
+trivial, and from the MathML output, it is easy to extract the identifiers.
+Second, the articles are already annotated with mark-up. Particularly, hyperlinks
+to other articles within Wikipedia are of interest as they typically wrap
+around any number of words and indicate that these in combination are relevant
+in the given context or (respectively) sentence.
 
@@
 framework, one can make use of a parallel processing system to speed up
-computation. Therefore we implement the proposed strategy with the
-Stratosphere system based on the PACT programming model \cite{Alexandrov2010}.
+computation. We implement the proposed strategy with the Stratosphere system
+\cite{stratosphere}. It is based on the PACT programming model
+\cite{Alexandrov2010}, which enables us to rapidly generate a large amount of
+definition relation candidates with only minimal implementation overhead for
+the parallelization.
 
@@
 that are matched by a pattern. The patterns being used to discover description
-terms are depicted in table \ref{tpatterns}. Due to the fact that we already
-tokenized an annotated the articles in a previous step in the MLP system, we
+terms are depicted in Table \ref{tpatterns}. Due to the fact that we already
+tokenized and annotated the articles in a previous step in the MLP system, we
 can make use of POS tags here as well.
@@
 Note, determiners not only contain articles, but also quantifiers and
-distributives. The last pattern in table \ref{tpatterns} contains `*/DT'. This
-is a shorthand for every word, that has the POS tag `DT' (determiner), will be
-matched at this position. Otherwise this pattern whould be rather large, as it
-needs to contain every possible determiner. \emph{identifer} as well as
-\emph{description} are placholders, that mark the positions of the entities
-from a possible defintion relation.
+distributives. The last pattern in Table \ref{tpatterns} contains `*/DT'. This
+is a shorthand for every word, that has the POS tag `DT' (determiner).
+Otherwise this pattern would be rather large, as it needs to contain every
+possible determiner. \emph{IDENTIFIER} as well as \emph{DESCRIPTION} are
+place-holder, that mark the positions of the entities from a possible definition
+relation.
 
@@
 
-Extracting relevant identifiers from the article rely on the assumption that
+Extracting relevant identifiers from the article relies on the assumption that
 the author will use \texttt{<math/>} tags for all formulae. This said, a
-formula that is written within the text cannot be recognized, and therefore
-can never be extracted by our system.
+formula that is written in the running text cannot be recognized, and therefore,
+cannot be extracted by our system.
 
 The fact that we can estimate all relevant identifiers for an article (see
-Section \ref{ir}), combined with some common assumptions about the relation,
+Section \ref{ir}), combined with some common assumptions about definition relations,
 can be exploited to largely reduce the set of candidates that need to be
@@
 relation co-occur in the same sentence. In other words, if we want to retrieve
-the description for an identifier, in the set of sentences that contain the
-identifier the description should be found. Thus said, any other sentences can
+the description for an identifier, only sentences containing the identifier
+could include the definition relation. Having said this, any other sentences can
 be ignored. Furthermore, we assume that it is more likely to find the
 description in first sentences than in the latter. This is based on the idea
-that authors introduce the meaning of an identifer and than subsequent use it,
-without necessarily repeating the meaning.
+that authors introduce the meaning of an identifier and than subsequent use the
+identifier, without necessarily repeating its definition.
 
-Another assumption can be made about the lexical class of the description we
-want to rank. They are nouns or even noun phrases (e.g., \texttt{`the
+Another assumption can be made about the lexical class of the definition relation we
+want to rank. The descriptions are nouns or even noun phrases (e.g., \texttt{`the
 effective spring constant k'} or \texttt{`mass m of something'}). We discard
 all other words (according to their POS tag) except noun phrases and Wikipedia
-hyperlinks. These are the candidates for a definition relation. Noun phrases
-and hyperlinks may consist of multiple words. For all intents and purposes, it
-is not necessary to threat noun phrases and hyperlinks as a set of words, and
-therefore, will be threated subsequently as if they were one. This is
+hyper-links. These are the candidates descriptions for a definition relation. Noun phrases
+and hyper-links may consist of multiple words. For all intents and purposes, it
+is not necessary to threat noun phrases and hyper-links as a set of words, and
+therefore, they will be treated subsequently as if they were one. This is
 important, due to the fact that the overall ranking will be greatly influenced
-by the distance of candidates to the position of the identifer.
+by the distance of candidates to the position of the identifier.
 
@@
 accuracy of retrieving the correct set of identifiers. First of all, people
-tend to incorrectly use \emph{texvc} when trying to create formulas. For example,
+tend to use \emph{texvc} only as a typesetting language and neglect its semantic capabilities. For example,
 \texttt{\textbackslash text\{log\}} is more often used than the correct
@@
 
-We took a very conservative approach and preprocessed all formulas.
-The \TeX~command \texttt{\textbackslash text\{\}} blocks along with subscriptions containing
-more than a single character will be removed before analysis. Superscripts
-will also be ignored in terms of being a part of the identifier. Moreover, we
-created a comprehensive blacklist to improve the results further. Short words
-like `a', `A', and `I', which are very common in the English language, could be
-easily matched by our processor in the surrounding text if a formula contains
-one of those identifiers, and therefore, will also be blacklisted. Additionally,
-we blacklist common mathematical operators, constants, and functions.
+We took a very conservative approach and preprocessed all formulas. The
+\TeX~command \texttt{\textbackslash text\{\}} blocks along with subscriptions
+containing more than a single character will be removed before analysis.
+Superscripts will also be ignored in terms of being a part of the identifier.
+Moreover, we created a comprehensive blacklist to improve the results further.
+Identifier like `a', `A', and `I', which are also very common in the English
+language, could be easily matched by our processor in the surrounding text,
+and therefore, will also be blacklisted. Additionally, we blacklist common
+mathematical operators, constants, and functions.
 
-Because we do not have any annotated test corpora, evaluation has to be
-performed by hand. Therefore, we took a sample of 30 random articles and
-counted all matches. The resulting estimates for the identifier retrieval
-performance are: \emph{Recall: 0.99} and \emph{Precision: 0.86}, which satisfy
-our information needs, as we are mostly interested in recall at this stage.
+We took a sample of 30 random articles and counted all matches by hand. The
+resulting estimates for the identifier retrieval performance are
+\emph{Recall: 0.99} and \emph{Precision: 0.86}, which satisfy our information
+needs, as we are mostly interested in recall at this stage.
 
@@
 
-There is no gold standard evaluation datasets to measure the performance of
-identifier definition extractors. Thus we created on on our own dataset. This
-is a very time consuming job. At the moment, the dataset only contains two
-large articles (revision ids are included) with around 100 identifier
-definitions. This dataset is also available on the project repository.
+At the start of our project there were no gold standard datasets available to
+measure the performance of identifier definition extractors. Thus, we created
+one on our own. This is a very time consuming job. At the moment, the dataset
+only contains two large articles (revision ids are included) with around 100
+identifier definitions. This dataset is also available on the project
+repository.
 
@@
 In that situation, other methods, especially supervised ones, are preferred.
-Unfortunately, many of them require a labeled test corpus to measure the
+Unfortunately, many of them require a labelled test corpus to measure the
 performance of a classifier that could be trained with our generated data.

### 2014-07-05 paper.tex
added bibtex and vspace before bibliography
Refernces fix
* add required refernces (ref1, ref2)
* fix stratosphere reference
Fixup: wrong explaination of the ranking formulae

@@
 \citeauthor{Yokoi} \cite{Yokoi} trained a \emph{support vector machine} to extract
-natural language descriptions for mathematical expressions.
+natural language descriptions for mathematical expressions. Furhter work in this field was
+done by \cite{ref2} and \cite{ref1}.
 
@@
 +\beta{R}_{\sigma_\mathrm s}(n)
-+\gamma\,\mathrm{tf}(t,s)}{\alpha+\beta+\gamma} \mapsto [0,1].
++\gamma\,\mathrm{tf}(t,d)}{\alpha+\beta+\gamma} \mapsto [0,1].
 \end{equation}
@@
 The weighted sum depends on the distance $\Delta$ (amount of word tokens) between
-identifier and the description term $t$, the sentence number $n$ counting
-(from the beginning of the article) all sentences containing the identifier,
-and the term frequency $\mathrm{tf}(t,s)$ in the set of sentences $s$. The
+identifier and the description term $t$, the distance between formula and sentence $n$,
+and the term frequency $\mathrm{tf}(t,d)$. The
 distance was normalized with $R_\sigma(\Delta) = \exp\left[ -\frac{1}{2}
@@
 seen such as `Hamiltonian'. These are both valid definitions for identifiers.
-As the influence of $\mathrm{tf}(t,s)$ on the sensitivity of the overall ranking
+As the influence of $\mathrm{tf}(t,d)$ on the sensitivity of the overall ranking
 (\ref{eq:rating}) seems to be very high, we reduce the impact with the tuning
@@
 
-\begingroup
+\let\oldclearpage\clearpage
 \let\clearpage\relax
+\vspace{10 mm}
 \bibliography{mlp-papers}
-\endgroup
+\let\clearpage\oldclearpage
+
+% \let\standardclearpage\clearpage
+% \begingroup
+% \let\clearpage\relax
+% \endgroup
 
